---
title: |
  | Detroit Council Districts 1 - 7 Regression Model Comparison:
  | R | SPSS&nbsp;&nbsp;&nbsp;&nbsp;2023 - 2025 Sales
author: |
  | Elliot Dean - Program Analyst IV  
  | Office of the Chief Financial Officer | Office of the Assessor
  | City of Detroit
date: "`r sub('^0', '', format(Sys.Date(), '%d %B %Y'))`"
geometry: margin = 1.5cm
output:
  pdf_document:
    latex_engine: xelatex
    toc: true
    toc_depth: 3
mainfont: Arial
spacing: 1.5
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, error = FALSE, echo = FALSE)
```

```{r lib_load}
library(tidyverse)   # data wrangling & ggplot2
library(cowplot)     # Streamlined Plot Theme and Plot Annotations for 'ggplot2'
library(patchwork)   # The Composer of Plots
library(janitor)     # functions for examining and cleaning dirty data
library(kableExtra)  # simple table generator
library(here)        # Find files within a project
library(olsrr)       # package for p-value backward elimination
library(broom)       # summarizes key information about models
library(rlang)       # functions for Base Types and Core R and 'Tidyverse' Features
library(car)         # VIF, influence measures helpers
library(scales)      # provides the internal scaling infrastructure used by ggplot
library(effects)     # Effect Displays for Linear, Generalized Linear, and Other Models
```

```{r fun, results='hide'}
codf     <- readRDS(file = here("rmds/modfunctions", "codfunctions.rds"))
codform  <- readRDS(file = here("rmds", "codformulas.rds"))
codf_env <- list2env(codf, parent = globalenv())
for (nm in names(codf)) {environment(codf[[nm]]) <- codf_env}
list2env(codf, envir = codf_env)
source(here("rmds/cod2325/districtruns/d2325f", "spss_cod_district-7_23-25f.R"), local = knitr::knit_global())
```

```{r loadmodsdata}
load(here("rmds/cod2325/districtruns/dmods", "spss_cod_district-1_23-25_June5.RData"))
load(here("rmds/cod2325/districtruns/dmods", "spss_cod_district-2_23-25_June5.RData"))
load(here("rmds/cod2325/districtruns/dmods", "spss_cod_district-3_23-25_June5.RData"))
load(here("rmds/cod2325/districtruns/dmods", "spss_cod_district-4_23-25_June2.RData"))
load(here("rmds/cod2325/districtruns/dmods", "spss_cod_district-5_23-25_June2.RData"))
load(here("rmds/cod2325/districtruns/dmods", "spss_cod_district-6_23-25_June2.RData"))
load(here("rmds/cod2325/districtruns/dmods", "spss_cod_district-7_23-25_June2.RData"))
dfspss <- read_csv(here("data/spss", "spssmods.csv"))
spsdf  <- read_csv(here("data/spss", "spssmods1.csv"))
dfsps  <- read_csv(here("data/spss", "spsscos.csv"))
```

```{r heatmap_set_r}
r_models <- list(
  D1 = list(M1 = d1_mod_time_23_25_June5, M2 = d1_mod_ln_price_no_ecf_23_25_June5,
            M3 = d1_mod_base_23_25_June5, M4 = d1_mod_base_inliers_23_25_June5,
            M5 = d1_mod_tasp_23_25_June5),
  D2 = list(M1 = d2_mod_time_23_25_June5, M2 = d2_mod_ln_price_no_ecf_23_25_June5,
            M3 = d2_mod_base_23_25_June5, M4 = d2_mod_base_inliers_23_25_June5,
            M5 = d2_mod_tasp_23_25_June5),
  D3 = list(M1 = d3_mod_time_23_25_June5, M2 = d3_mod_ln_price_no_ecf_23_25_June5,
            M3 = d3_mod_base_23_25_June5, M4 = d3_mod_base_inliers_23_25_June5,
            M5 = d3_mod_tasp_23_25_June5),
  D4 = list(M1 = d4_mod_time_23_25_June2, M2 = d4_mod_ln_price_no_ecf_23_25_June2,
            M3 = d4_mod_base_23_25_June2, M4 = d4_mod_base_inliers_23_25_June2,
            M5 = d4_mod_tasp_23_25_June2),
  D5 = list(M1 = d5_mod_time_23_25_June2, M2 = d5_mod_ln_price_no_ecf_23_25_June2,
            M3 = d5_mod_base_23_25_June2, M4 = d5_mod_base_inliers_23_25_June2,
            M5 = d5_mod_tasp_23_25_June2),
  D6 = list(M1 = d6_mod_time_23_25_June2, M2 = d6_mod_ln_price_no_ecf_23_25_June2,
            M3 = d6_mod_base_23_25_June2, M4 = d6_mod_base_inliers_23_25_June2,
            M5 = d6_mod_tasp_23_25_June2),
  D7 = list(M1 = d7_mod_time_23_25_June2, M2 = d7_mod_ln_price_no_ecf_23_25_June2,
            M3 = d7_mod_base_23_25_June2, M4 = d7_mod_base_inliers_23_25_June2,
            M5 = d7_mod_tasp_23_25_June2)
)

extract_lm_metrics <- function(mod) {
  sm <- summary(mod)
  f  <- sm$fstatistic
  tibble(
    r = sqrt(max(sm$r.squared, 0)),   r_squared = sm$r.squared,
    adj_r_squared = sm$adj.r.squared, see = sm$sigma,
    f_stat = if (!is.null(f)) unname(f["value"]) else NA_real_
  )
}

r_fit <- imap_dfr(
  r_models,
  \(district_models, district) {
    imap_dfr(
      district_models,
      \(mod, model) {
        extract_lm_metrics(mod) |> mutate(district = district, model = model, .before = 1)
      }
    )
  }
)

extract_ratio_metrics <- function(mod, response_is_logged = TRUE, correction_factor = 1) {
  mf     <- model.frame(mod)
  actual <- model.response(mf)
  pred   <- fitted(mod)

  if (response_is_logged) {
    actual <- exp(actual)
    pred   <- exp(pred) * correction_factor
  }

  ok           <- is.finite(actual) & is.finite(pred) & actual > 0 & pred > 0
  actual       <- actual[ok]
  pred         <- pred[ok]
  ratio        <- pred / actual
  median_ratio <- median(ratio)
  mean_ratio   <- mean(ratio)
  wmean_ratio  <- sum(pred) / sum(actual)
  COD          <- mean(abs(ratio - median_ratio)) / median_ratio
  PRD          <- mean_ratio / wmean_ratio
  value_proxy  <- 0.5 * (actual + pred / median_ratio)

  prb_df <- tibble(y = ratio / median_ratio - 1, x = log2(value_proxy)) |>
    filter(is.finite(x), is.finite(y))
  
  ok_ratio <- is.finite(ratio)
  r_ok     <- ratio[ok_ratio]
  n_ok     <- length(r_ok)
  
  within_med_n <- function(pct) {
    lo <- (1 - pct/100) * median_ratio
    hi <- (1 + pct/100) * median_ratio
    sum(r_ok >= lo & r_ok <= hi, na.rm = TRUE)
  }
  
  w10_n <- within_med_n(10)
  w20_n <- within_med_n(20)
  w50_n <- within_med_n(50)
  w10_pct <- if (n_ok > 0) 100 * w10_n / n_ok else NA_real_
  w20_pct <- if (n_ok > 0) 100 * w20_n / n_ok else NA_real_
  w50_pct <- if (n_ok > 0) 100 * w50_n / n_ok else NA_real_

  PRB <- if (nrow(prb_df) >= 3) {coef(lm(y ~ x, data = prb_df))[["x"]]} else {NA_real_}
  tibble(median_ratio = median_ratio, COD = COD, PRD = PRD, PRB = PRB, within_10_pct = w10_pct,
    within_20_pct = w20_pct, within_50_pct = w50_pct)
}

r_ratio <- imap_dfr(
  r_models,
  \(district_models, district) {
    imap_dfr(
      district_models,
      \(mod, model) {
        extract_ratio_metrics(mod, response_is_logged = TRUE) |>
          mutate(district = district, model = model, .before = 1)
      }
    )
  }
)

r_metrics <- left_join(r_fit, r_ratio, by = c("district", "model")) |> dplyr::mutate(source = "R")

spss_metrics <- spsdf |> mutate(district = paste0("D", codd), model = paste0("M", model)) |> 
  rename(r_squared = r2, adj_r_squared = adj_r2) |> dplyr::mutate(source = "SPSS")
```

```{r hm2}
metrics_to_compare <- c(
  "r_squared", "adj_r_squared", "see", "f_stat", "median_ratio", "COD", "PRD", "PRB",
  "within_10_pct", "within_20_pct", "within_50_pct"
)

r_long <- r_metrics |> select(district, model, all_of(metrics_to_compare)) |>
  pivot_longer(cols = all_of(metrics_to_compare), names_to = "metric", values_to = "r_value")

spss_long <- spss_metrics |> select(district, model, all_of(metrics_to_compare)) |>
  pivot_longer(cols = all_of(metrics_to_compare), names_to = "metric", values_to = "spss_value")

comparison <- r_long |> inner_join(spss_long, by = c("district", "model", "metric"))

metric_tolerances <- tribble(
  ~metric,          ~tolerance,
  "r_squared",       0.005,
  "adj_r_squared",   0.005,
  "see",             0.010,
  "f_stat",          1.000,
  "median_ratio",    0.005,
  "COD",             0.500,
  "PRD",             0.005,
  "PRB",             0.005,
  "within_10_pct",   0.100,
  "within_20_pct",   0.100,
  "within_50_pct",   0.100
)

comparison <- comparison |> left_join(metric_tolerances, by = "metric") |>
  mutate(difference = r_value - spss_value,
    abs_difference = abs(difference), discrepancy_score = abs_difference / tolerance)

comparison <- comparison |>
  mutate(district_num = readr::parse_number(district),
    model_num = readr::parse_number(model),
    model_id = paste0("District ", district_num, " – Model ", model_num)) |>
  arrange(district_num, model_num) |>
  mutate(model_id = factor(model_id, levels = rev(unique(model_id))))

metric_labels <- c(
  r_squared     = "R²",           adj_r_squared = "Adj. R²",    see = "SEE", f_stat = "F",
  median_ratio  = "Median Ratio", COD = "COD",                  PRD = "PRD", PRB    = "PRB",
  within_10_pct = "Within 10%",   within_20_pct = "Within 20%", within_50_pct = "Within 50%"
)

comparison <- comparison |>
  mutate(metric_label = factor(metric_labels[metric], levels = unname(metric_labels)))
#
metric_rules <- tibble::tribble(
  ~metric,           ~rule,             ~target,
  "r",               "higher_better",    NA_real_,
  "r_squared",       "higher_better",    NA_real_,
  "adj_r_squared",   "higher_better",    NA_real_,
  "see",             "lower_better",     NA_real_,
  "rmse",            "lower_better",     NA_real_,
  "mae",             "lower_better",     NA_real_,
  "f_stat",          "higher_better",    NA_real_,   # optional
  "median_ratio",    "closer_to_target", 1.00,
  "COD",             "lower_better",     NA_real_,
  "PRD",             "closer_to_target", 1.00,
  "PRB",             "closer_to_target", 0.00,
  "within_10_pct",   "higher_better",    NA_real_,
  "within_20_pct",   "higher_better",    NA_real_,
  "within_50_pct",   "higher_better",    NA_real_
)

strength_df <- comparison |> left_join(metric_rules, by = "metric") |>
  mutate(
    raw_strength = case_when(
      rule == "higher_better" ~ r_value - spss_value,
      rule == "lower_better"  ~ spss_value - r_value,
      rule == "closer_to_target" ~
        abs(spss_value - target) - abs(r_value - target), TRUE ~ NA_real_
    )
  )

strength_df <- strength_df |> group_by(metric) |>
  mutate(
    max_abs_strength = max(abs(raw_strength), na.rm = TRUE),
    strength_scaled = case_when(
      is.na(raw_strength) ~ NA_real_,
      max_abs_strength == 0 ~ 0, TRUE ~ raw_strength / max_abs_strength
    )
  ) |> ungroup()

metric_order <- c(
  "R", "R²", "Adj. R²", "SEE", "RMSE", "MAE", "F",
  "Median Ratio", "COD", "PRD", "PRB", "Within 10%", "Within 20%", "Within 50%"
)

strength_df <- strength_df |>
  mutate(
    metric_label = metric_labels[metric],
    metric_label = factor(metric_label, levels = metric_order)
  )

strength_df <- strength_df |>
  mutate(
    diff_label = case_when(
      metric %in% c("COD") ~ sprintf("%+.2f", r_value - spss_value),
      metric %in% c("within_10_pct", "within_20_pct", "within_50_pct") ~
        sprintf("%+.2f", r_value - spss_value), TRUE ~ sprintf("%+.4f", r_value - spss_value)
    )
  )

metric_labels2 <- c(
  r              = "R",
  r_squared      = "R²",
  adj_r_squared  = "Adj. R²",
  see            = "SEE",
  rmse           = "RMSE",
  mae            = "MAE",
  f_stat         = "F",
  median_ratio   = "Median Ratio",
  COD            = "COD",
  PRD            = "PRD",
  PRB            = "PRB",
  within_10_pct  = "Within 10%",
  within_20_pct  = "Within 20%",
  within_50_pct  = "Within 50%"
)

strength_df2 <- strength_df |>
  mutate(
    district_num2 = readr::parse_number(as.character(district)),
    model_num2    = readr::parse_number(as.character(model)),
    row_id = paste0("District ", district_num2, " – Model ", model_num2),
    row_id = factor(row_id, levels = rev(unique(row_id[order(district_num2, model_num2)]))),
    metric_label2 = factor(metric_labels2[metric],
      levels = unname(metric_labels2[metric_labels2 %in% metric_labels2]))
  )
```

```{r heatmapplot, eval=FALSE}
heatmap_plot <- ggplot(
  comparison, aes(x = metric_label, y = model_id, fill = discrepancy_score)
) + geom_tile(color = "white", linewidth = 0.7) +
  scale_fill_gradientn(
    colours = c("#1a9850", "#91cf60", "#fee08b", "#fc8d59", "#d73027"),
    values = scales::rescale(c(0, 0.5, 1, 2, 3)), limits = c(0, 3), oob = scales::squish,
    name = "Difference /\nTolerance"
  ) +
  labs(
    title = "SPSS vs R Model Reproducibility",
    subtitle = "Comparison across 7 districts and 5 model specifications",
    x = NULL, y = NULL
  ) + theme_minimal(base_size = 11) +
  theme(
    panel.grid = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1, face = "bold"),
    axis.text.y = element_text(size = 8), plot.title = element_text(face = "bold", size = 16),
    legend.position = "right"
  )
heatmap_plot
```

---

## Introduction

This report contains multiple comparisons of linear regression models prepared in SPSS, and R.
In order to assure model reproducibility, datasets and methods utilized in SPSS were implemented in
R as closely as possible. Due to the nature of stepwise backward elimination however, exact
thresholds for predictor removal could not be fine tuned for every model designed for all 7
districts. Overall, models created in R using the same parameters as those in SPSS had
exceptionally close performance metrics.

These results demonstrate that the current linear regression model approach is viable across
platforms. 

---

## Coefficient-Concordance

The coefficients represented on the graph below indicate: 1) The direction, and 2) The impact of a
predictor on the dependent variable (for these models home value). Values close to zero have a 
negligible impact on sale price. Strong negative values correlate with decreased sale price
(e.g., unsound condition properties), high positive values correlated with increased sale price
(e.g., properties in good condition).

As SPSS and R coefficients overlap significantly, they produce a line with slope = 1 when plotted
on opposite axes.

```{r coefsetup}
r_coefs <- purrr::imap_dfr(
  r_models,
  \(district_models, district) {
    purrr::imap_dfr(
      district_models,
      \(mod, model) {
        tibble(predictor = names(stats::coef(mod)),
          r_coef = unname(stats::coef(mod)), district = district, model = model, source = "R")
      }
    )
  }
)

dfsps$term <- tolower(dfsps$term)
spss_coefs <- dfsps |> dplyr::rename(spss_coef = b) |> 
  dplyr::mutate(
    predictor = dplyr::case_when(term == "(constant)" ~ "(Intercept)", TRUE ~ term),
    district = paste0("D", codd), model = paste0("M", model), source = "SPSS"
  )

coef_compare <- r_coefs |>
  dplyr::inner_join(spss_coefs, by = c("district", "model", "predictor")) |>
  dplyr::mutate(coef_diff = r_coef - spss_coef, abs_diff = abs(coef_diff))
```

```{r coefplot1, fig.dim=c(8,5)}
# 1. Establish overall coefficient limits
coef_min <- min(c(coef_compare$spss_coef, coef_compare$r_coef), na.rm = TRUE)
coef_max <- max(c(coef_compare$spss_coef, coef_compare$r_coef), na.rm = TRUE)
# 2.Split coefficients into two ranges Use midpoint so each SPSS/R coefficient pair stays together
coef_compare_plot <- coef_compare |>
  mutate(
    coef_midpoint = (spss_coef + r_coef) / 2,
    coef_range = case_when(coef_midpoint < 6  ~ "Below 6", coef_midpoint >= 6 ~ "6 and Above")
  )
# 3. LOWER-RANGE PANEL
p_low <- coef_compare_plot |> filter(coef_range == "Below 6") |>
  ggplot(aes(x = spss_coef, y = r_coef)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", linewidth = 0.8, color = "black") +
  geom_point(shape = 21, size = 3, stroke = 0.7,
    fill = "#0072B2",       # R
    color = "#D55E00",      # SPSS
    alpha = 0.60
  ) +
  coord_equal(xlim = c(coef_min, 6), ylim = c(coef_min, 6), expand = FALSE) +
  labs(title = "Coefficient < 6", x = "SPSS Coefficient", y = "R Coefficient") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank(),
    plot.title = element_text(
      face = "bold", hjust = 0.5), axis.title = element_text(face = "bold")
  )
# 4. UPPER-RANGE PANEL
p_high <- coef_compare_plot |> filter(coef_range == "6 and Above") |>
  ggplot(aes(x = spss_coef, y = r_coef)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", linewidth = 0.8, color = "black") +
  geom_point(shape = 21, size = 3, stroke = 0.7, fill = "#0072B2",       
    color = "#D55E00", alpha = 0.60) +
  coord_equal(xlim = c(6, coef_max), ylim = c(6, coef_max), expand = FALSE) +
  labs(title = "Coefficient >= 6", x = "SPSS Coefficient", y = "R Coefficient") +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold", hjust = 0.5), axis.title = element_text(face = "bold")
  )

p_coef_split <-
  p_low + p_high + plot_layout(widths = c(1.5, 1)) +
  plot_annotation(title = "SPSS | R Coefficient Concordance",
    subtitle = paste(
      "Points on the dashed 1:1 line indicate equivalent", "coefficient estimates"
    ), theme = theme(plot.title = element_text(face = "bold", size = 15))
  )

p_coef_split
```

---

## Heatmap Overview

The following figure highlights the similarity (or difference) of a metric from an R model
compared to the SPSS created equivalent.

```{r hmapplot, fig.dim=c(8,9), eval=FALSE}
comparison |> mutate(model_label = paste0("Model ", model_num)) |>
  ggplot(aes(x = metric_label, y = model_label, fill = discrepancy_score)) +
  geom_tile(color = "white", linewidth = 0.7) +
  facet_wrap(~ district, ncol = 2) + # geom_text(aes(label = difference_label), size = 2.4) +
  scale_fill_gradientn(
    colours = c("#1a9850", "#91cf60", "#fee08b", "#fc8d59", "#d73027"),
    values = scales::rescale(c(0, 0.5, 1, 2, 3)), limits = c(0, 3),
    oob = scales::squish, name = "R − SPSS" # |R − SPSS| /\nTolerance
  ) +
  labs(
    title = "SPSS versus R Model Reproducibility", subtitle = "7 Districts × 5 AVM Specifications",
    x = NULL, y = NULL) + theme_minimal(base_size = 11) +
  theme(
    panel.grid = element_blank(), strip.text = element_text(face = "bold", size = 11),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8)
  )
```

```{r hmapplot2, fig.dim=c(8,10)}
p_strength <- ggplot(strength_df2, aes(x = metric_label2, y = row_id, fill = strength_scaled)) +
  geom_tile(color = "white", linewidth = 0.7) +
  geom_text(aes(label = diff_label), size = 2.75) +
  scale_fill_gradient2(low = "#D55E00",     # worse for R
    mid = "white", high = "#0072B2",    # better for R
    midpoint = 0, limits = c(-1, 1), oob = scales::squish, name = "R relative\nto SPSS"
  ) +
  labs(
    title = "R | SPSS Performance Metric Comparison", # Strength/Weakness Relative to SPSS
    subtitle = "Blue = R higher; orange = R lower; cell text shows R − SPSS", #stronger, weaker
    x = NULL, y = NULL
  ) + theme_minimal(base_size = 11) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, face = "bold"),
    axis.text.y = element_text(size = 9, face = "bold"),
    plot.title = element_text(face = "bold", size = 15), legend.position = "bottom"
  )
p_strength
```

```{r hmapplot3, fig.dim=c(8,10), eval=FALSE}
strength_df_facet <- strength_df |>
  mutate(
    district_label = paste0("District ", district_num),
    model_label = factor(paste0("Model ", model_num),
      levels = rev(paste0("Model ", sort(unique(model_num)))))
  )

p_strength_facet <- ggplot(
  strength_df_facet, aes(x = metric_label, y = model_label,  fill = strength_scaled)) +
  geom_tile(color = "white", linewidth = 0.7) +
  # geom_text(aes(label = diff_label), size = 2.4) +
  facet_wrap(~ district_label, ncol = 2) +
  scale_fill_gradient2(low = "#D55E00", mid = "white",  high = "#0072B2", midpoint = 0,
    limits = c(-1, 1), oob = scales::squish, name = "R relative\nto SPSS") +
  labs(title = "R Strength/Weakness Relative to SPSS",
    subtitle = "Blue = R stronger; orange = R weaker", x = NULL, y = NULL) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid = element_blank(), strip.text = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 9, face = "bold"),
    axis.text.y = element_text(face = "bold"),
    plot.title = element_text(face = "bold", size = 16), panel.spacing = grid::unit(0.8, "cm")
  )
p_strength_facet
```

---

## Metric Comparison

The following dumbbell plots identify overlap (or disparity) of specific metric values from R and
SPSS models, arranged by district.

```{r db1}
prep_dumbbell_data <- function(data, metric_name) {
  plot_df <- data |> dplyr::filter(metric == metric_name) |>
    dplyr::mutate(
      district_num = readr::parse_number(as.character(district)),
      model_num    = readr::parse_number(as.character(model)),
      district_lab = paste0("District ", district_num),
      model_lab    = paste0("Model ", model_num),
      difference   = r_value - spss_value,
      midpoint     = (r_value + spss_value) / 2
    ) |> dplyr::arrange(district_num, model_num) |>
    dplyr::mutate(
      district_lab = factor(
        district_lab, levels = paste0("District ", sort(unique(district_num)))
      ),    # reverse so Model 1 is at the top
      model_lab = factor(model_lab, levels = rev(paste0("Model ", sort(unique(model_num))))),
      row_group = dplyr::case_when(
        district_num %in% c(1, 2) ~ 1, district_num %in% c(3, 4) ~ 2,
        district_num %in% c(5, 6) ~ 3, district_num %in% c(7)    ~ 4
      )
    )
  
  point_df <- plot_df |>
    dplyr::select(district_num, district_lab, row_group, model_lab, spss_value, r_value) |>
    tidyr::pivot_longer(
      cols = c(spss_value, r_value), names_to = "Software", values_to = "value"
    ) |>
    dplyr::mutate(
      Software = dplyr::case_when(
        Software == "spss_value" ~ "SPSS", Software == "r_value" ~ "R", TRUE ~ Software
      )
    )
  list(plot_df = plot_df, point_df = point_df)
}

make_dumbbell_column <- function(
    plot_df, point_df, districts_keep, side = c("left", "right")) {
  side <- match.arg(side)
  plot_df_sub <- plot_df |>
    dplyr::filter(district_num %in% districts_keep) |>
    dplyr::mutate(facet_row = factor(row_group, levels = 1:4))

  point_df_sub <- point_df |> dplyr::filter(district_num %in% districts_keep) |>
    dplyr::mutate(facet_row = factor(row_group, levels = 1:4))

  facet_labels <- if (side == "left") {
    c(`1` = "District 1", `2` = "District 3", `3` = "District 5", `4` = "District 7")
  } else {
    c(`1` = "District 2", `2` = "District 4", `3` = "District 6", `4` = "")   # blank fourth row
  }

  ggplot() +
    geom_segment(
      data = plot_df_sub, aes(x = spss_value, xend = r_value, y = model_lab, yend = model_lab),
      linewidth = 1, color = "grey60"
    ) +
    geom_point(data = point_df_sub, aes(x = value, y = model_lab, color = Software), size = 3.5) +
    facet_wrap(
      ~ facet_row, ncol = 1, drop = FALSE, labeller = labeller(facet_row = facet_labels)
    ) + scale_color_manual(values = c("SPSS" = "#D55E00", "R" = "#0072B2")) +
    labs(x = NULL, y = NULL, color = NULL) + theme_minimal(base_size = 12) +
    theme(
      panel.grid.major.y = element_blank(),
      panel.grid.minor = element_blank(), panel.spacing.y = grid::unit(0.8, "cm"),
      strip.text = element_text(face = "bold", size = 11), strip.background = element_blank(),
      axis.text.y = element_blank(), axis.ticks.y = element_blank(), legend.position = "none"
    )
}

make_center_model_axis <- function() {
  center_df <- expand.grid(
    row_group = factor(1:4),
    model_lab = factor(rev(paste0("Model ", 1:5)), levels = rev(paste0("Model ", 1:5)))
  )
  
  ggplot(center_df, aes(x = 1, y = model_lab, label = model_lab)) +
    geom_text(fontface = "bold", size = 4) + facet_wrap(~ row_group, ncol = 1) +
    scale_x_continuous(limits = c(0.8, 1.2)) + labs(x = NULL, y = NULL) +
    theme_void(base_size = 12) +
    theme(panel.spacing.y = grid::unit(0.9, "cm"), strip.text = element_blank())
}

make_dumbbell_center_axis <- function(data, metric_name, metric_title = metric_name) {
  dat      <- prep_dumbbell_data(data, metric_name)
  plot_df  <- dat$plot_df
  point_df <- dat$point_df
  
  p_left <- make_dumbbell_column(
    plot_df = plot_df, point_df = point_df, districts_keep = c(1, 3, 5, 7), side = "left"
  )
  
  p_mid <- make_center_model_axis()
  
  p_right <- make_dumbbell_column(
    plot_df = plot_df, point_df = point_df, districts_keep = c(2, 4, 6), side = "right"
  )
  
  legend_plot <- ggplot(point_df, aes(x = value, y = model_lab, color = Software)) + geom_point() +
    scale_color_manual(values = c("SPSS" = "#D55E00", "R" = "#0072B2")) +
    theme_void() + theme(legend.position = "top", axis.text.x = element_text(face = "bold"))
  
  leg <- cowplot::get_legend(legend_plot)
  
  body <- cowplot::plot_grid(
    p_left, p_mid, p_right, nrow = 1, rel_widths = c(1, 0.18, 1), align = "h", axis = "tb"
  )
  
  title <- ggdraw() +
    draw_label(
      paste0(metric_title, ": SPSS | R"), x = 0, hjust = 0, fontface = "bold", size = 15
    )
  
  subtitle <- ggdraw() +
    draw_label(
      "Each line connects the SPSS and R result for the same AVM", x = 0, hjust = 0, size = 12
    )
  
  xlab <- ggdraw() + draw_label(metric_title, size = 13, fontface = "bold")
  
  cowplot::plot_grid(
    title, subtitle, leg, body, xlab, ncol = 1, rel_heights = c(0.06, 0.05, 0.06, 1, 0.05)
  )
}
```

### Adjusted R²

```{r adjr2, fig.dim=c(8,8)}
p_adj_r2 <- make_dumbbell_center_axis(
  comparison, metric_name = "adj_r_squared", metric_title = "Adjusted R²"
)
p_adj_r2
```

### Median Ratio

```{r med, fig.dim=c(8,8)}
p_med <- make_dumbbell_center_axis(
  comparison, metric_name = "median_ratio", metric_title = "Median Ratio"
)
p_med
```

### Coefficient of Dispersion

```{r cod, fig.dim=c(8,8)}
p_cod <- make_dumbbell_center_axis(
  comparison, metric_name = "COD", metric_title = "Coefficient of Dispersion"
)
p_cod
```

### Price Related Differential

```{r prd, fig.dim=c(8,8)}
p_prd <- make_dumbbell_center_axis(
  comparison, metric_name = "PRD", metric_title = "Price Related Differential"
)
p_prd
```

### Price Related Bias

```{r prb, fig.dim=c(8,8)}
p_prb <- make_dumbbell_center_axis(
  comparison, metric_name = "PRB", metric_title = "Price Related Bias"
)
p_prb
```

```{r coefcheck, eval=FALSE}
coef_compare |> dplyr::count(district, model)
dplyr::anti_join(r_coefs, spss_coefs, by = c("district", "model", "predictor"))
```

```{r coefplot2, eval=FALSE}
coef_compare <- coef_compare |>
  dplyr::mutate(district_num = readr::parse_number(district),
    model_num = readr::parse_number(model),
    district = factor(district, levels = paste0("D", 1:7), labels = paste0("District ", 1:7)),
    model = factor(model, levels = paste0("M", 1:5), labels = paste0("Model ", 1:5))
  )

p_coef <- ggplot(coef_compare, aes(x = spss_coef, y = r_coef)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", linewidth = 0.7, color = "grey40") +
  geom_point(size = 2, alpha = 0.8) +
  facet_grid(rows = vars(district), cols = vars(model)) +  coord_equal() +
  labs(
    title = "SPSS vs R Coefficient Concordance",
    subtitle = "Seven districts × five AVM specifications",
    x = "SPSS Coefficient", y = "R Coefficient"
  ) +  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold"), plot.title = element_text(face = "bold", size = 16)
  )

p_coef
```


```{r iaao}

```
























