---
title: |
  | City Council Districts 1 - 7 Linear Regression Model Comparison:
  | R | SPSS    2023 - 2025 Sales
author: |
  | Elliot Dean - Program Analyst IV  
  | Office of the Chief Financial Officer | Office of the Assessor
  | City of Detroit
date: "`r sub('^0', '', format(Sys.Date(), '%d %B %Y'))`"
geometry: margin = 1.5cm
output:
  pdf_document:
    latex_engine: xelatex
    toc: true
mainfont: Arial
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, error = FALSE, echo = FALSE)
```

```{r lib_load}
library(tidyverse)   # data wrangling & ggplot2
library(janitor)     # functions for examining and cleaning dirty data
library(kableExtra)  # simple table generator
library(here)        # Find files within a project
library(olsrr)       # package for p-value backward elimination
library(broom)       # summarizes key information about models
library(rlang)       # functions for Base Types and Core R and 'Tidyverse' Features
library(car)         # VIF, influence measures helpers
library(scales)      # provides the internal scaling infrastructure used by ggplot
library(effects)     # Effect Displays for Linear, Generalized Linear, and Other Models
```

```{r fun, results='hide'}
codf     <- readRDS(file = here("rmds/modfunctions", "codfunctions.rds"))
codform  <- readRDS(file = here("rmds", "codformulas.rds"))
codf_env <- list2env(codf, parent = globalenv())
for (nm in names(codf)) {environment(codf[[nm]]) <- codf_env}
list2env(codf, envir = codf_env)
source(here("rmds/cod2325/districtruns/d2325f", "spss_cod_district-7_23-25f.R"), local = knitr::knit_global())
```

```{r loadmodsdata}
load(here("rmds/cod2325/districtruns/dmods", "spss_cod_district-1_23-25_June5.RData"))
load(here("rmds/cod2325/districtruns/dmods", "spss_cod_district-2_23-25_June5.RData"))
load(here("rmds/cod2325/districtruns/dmods", "spss_cod_district-3_23-25_June5.RData"))
load(here("rmds/cod2325/districtruns/dmods", "spss_cod_district-4_23-25_June2.RData"))
load(here("rmds/cod2325/districtruns/dmods", "spss_cod_district-5_23-25_June2.RData"))
load(here("rmds/cod2325/districtruns/dmods", "spss_cod_district-6_23-25_June2.RData"))
load(here("rmds/cod2325/districtruns/dmods", "spss_cod_district-7_23-25_June2.RData"))
dfspss <- read_csv(here("data/spss", "spssmods.csv"))
dfsps <- read_csv(here("data/spss", "spsscos.csv"))
```

---

## Model I - Logarithm of Sale Price per Square Foot (LN_SPPSF)

### Linear Regression Fit Statistics

```{r fit1m1}
rmt <- model_metrics_c(d1_mod_time_23_25_June5, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd1mod1") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 1 | LN\\_SPPSF Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit2m1}
rmt <- model_metrics_c(d2_mod_time_23_25_June5, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd2mod1") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 2 | LN\\_SPPSF Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit3m1}
rmt <- model_metrics_c(d3_mod_time_23_25_June5, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd3mod1") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 3 | LN\\_SPPSF Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit4m1}
rmt <- model_metrics_c(d4_mod_time_23_25_June2, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd4mod1") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 4 | LN\\_SPPSF Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit5m1}
rmt <- model_metrics_c(d5_mod_time_23_25_June2, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd5mod1") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 5 | LN\\_SPPSF Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit6m1}
rmt <- model_metrics_c(d6_mod_time_23_25_June2, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd6mod1") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 6 | LN\\_SPPSF Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit7m1}
rmt <- model_metrics_c(d7_mod_time_23_25_June2, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd7mod1") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 7 | LN\\_SPPSF Model Fit Statistics: SPSS vs R"
)
res$kable
```






---

## Model II - Logarithm of Sale Price (LN_PRICE) excluding ECFs

### Linear Regression Fit Statistics

```{r fit1m2}
rmt <- model_metrics_c(d1_mod_ln_price_no_ecf_23_25_June5, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd1mod2") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 1 | LN\\_PRICE (No ECFs) Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit2m2}
rmt <- model_metrics_c(d2_mod_ln_price_no_ecf_23_25_June5, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd2mod2") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 2 | LN\\_PRICE (No ECFs) Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit3m2}
rmt <- model_metrics_c(d3_mod_ln_price_no_ecf_23_25_June5, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd3mod2") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 3 | LN\\_PRICE (No ECFs) Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit4m2}
rmt <- model_metrics_c(d4_mod_ln_price_no_ecf_23_25_June2, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd4mod2") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 4 | LN\\_PRICE (No ECFs) Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit5m2}
rmt <- model_metrics_c(d5_mod_ln_price_no_ecf_23_25_June2, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd5mod2") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 5 | LN\\_PRICE (No ECFs) Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit6m2}
rmt <- model_metrics_c(d6_mod_ln_price_no_ecf_23_25_June2, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd6mod2") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 6 | LN\\_PRICE (No ECFs) Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit7m2}
rmt <- model_metrics_c(d7_mod_ln_price_no_ecf_23_25_June2, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd7mod2") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 7 | LN\\_PRICE (No ECFs) Model Fit Statistics: SPSS vs R"
)
res$kable
```







---

## Model III - Logarithm of Sale Price (LN_PRICE)

### Linear Regression Fit Statistics

```{r fit1m3}
rmt <- model_metrics_c(d1_mod_base_23_25_June5, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd1mod3") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 1 | LN\\_PRICE Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit2m3}
rmt <- model_metrics_c(d2_mod_base_23_25_June5, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd2mod3") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 2 | LN\\_PRICE Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit3m3}
rmt <- model_metrics_c(d3_mod_base_23_25_June5, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd3mod3") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 3 | LN\\_PRICE Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit4m3}
rmt <- model_metrics_c(d4_mod_base_23_25_June2, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd4mod3") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 4 | LN\\_PRICE Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit5m3}
rmt <- model_metrics_c(d5_mod_base_23_25_June2, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd5mod3") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 5 | LN\\_PRICE Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit6m3}
rmt <- model_metrics_c(d6_mod_base_23_25_June2, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd6mod3") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 6 | LN\\_PRICE Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit7m3}
rmt <- model_metrics_c(d7_mod_base_23_25_June2, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd7mod3") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 7 | LN\\_PRICE Model Fit Statistics: SPSS vs R"
)
res$kable
```







---

## Model IV - Logarithm of Sale Price (LN_PRICE) with Outlier Removal

### Linear Regression Fit Statistics

```{r fit1m4}
rmt <- model_metrics_c(d1_mod_base_inliers_23_25_June5, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd1mod4") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 1 | LN\\_PRICE (Inliers) Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit2m4}
rmt <- model_metrics_c(d2_mod_base_inliers_23_25_June5, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd2mod4") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 2 | LN\\_PRICE (Inliers) Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit3m4}
rmt <- model_metrics_c(d3_mod_base_inliers_23_25_June5, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd3mod4") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 3 | LN\\_PRICE (Inliers) Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit4m4}
rmt <- model_metrics_c(d4_mod_base_inliers_23_25_June2, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd4mod4") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 4 | LN\\_PRICE (Inliers) Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit5m4}
rmt <- model_metrics_c(d5_mod_base_inliers_23_25_June2, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd5mod4") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 5 | LN\\_PRICE (Inliers) Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit6m4}
rmt <- model_metrics_c(d6_mod_base_inliers_23_25_June2, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd6mod4") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 6 | LN\\_PRICE (Inliers) Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit7m4}
rmt <- model_metrics_c(d7_mod_base_inliers_23_25_June2, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd7mod4") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 7 | LN\\_PRICE (Inliers) Model Fit Statistics: SPSS vs R"
)
res$kable
```






---

## Model V - Logarithm of Time Adjusted Sale Price (LN_TASP)

### Linear Regression Fit Statistics

```{r fit1m5}
rmt <- model_metrics_c(d1_mod_tasp_23_25_June5, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd1mod5") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 1 | LN\\_TASP Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit2m5}
rmt <- model_metrics_c(d2_mod_tasp_23_25_June5, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd2mod5") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 2 | LN\\_TASP Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit3m5}
rmt <- model_metrics_c(d3_mod_tasp_23_25_June5, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd3mod5") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 3 | LN\\_TASP Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit4m5}
rmt <- model_metrics_c(d4_mod_tasp_23_25_June2, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd4mod5") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 4 | LN\\_TASP Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit5m5}
rmt <- model_metrics_c(d5_mod_tasp_23_25_June2, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd5mod5") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 5 | LN\\_TASP Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit6m5}
rmt <- model_metrics_c(d6_mod_tasp_23_25_June2, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd6mod5") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 6 | LN\\_TASP Model Fit Statistics: SPSS vs R"
)
res$kable
```

```{r fit7m5}
rmt <- model_metrics_c(d7_mod_tasp_23_25_June2, digits = 4)
smod1 <- dfspss |> filter(spssmod == "ccd7mod5") |> select(df_total:f_stat)
res <- make_metrics_kable(spss_df = smod1, r_df = rmt,
  caption = "District 7 | LN\\_TASP Model Fit Statistics: SPSS vs R"
)
res$kable
```

---

## Predictor Impact by District

### District 1

```{r blp1}
res_beta <- plot_avg_beta_lollipop_6model(
  d1_mod_time_23_25_June5, d1_mod_ln_price_no_ecf_23_25_June5,
  d1_mod_base_23_25_June5, d1_mod_base_inliers_23_25_June5,
  d1_mod_tasp_23_25_June5, d1_mod_alt_23_25_June5,
  model_names = c("M1", "M2", "M3", "M4", "M5", "M6"), top_n = 5,
  title = "District 1 | Average Standardized Beta across Models", include_ecf = FALSE
)
res_beta$plot
```

### District 2

```{r blp2}
res_beta <- plot_avg_beta_lollipop_6model(
  d2_mod_time_23_25_June5, d2_mod_ln_price_no_ecf_23_25_June5,
  d2_mod_base_23_25_June5, d2_mod_base_inliers_23_25_June5,
  d2_mod_tasp_23_25_June5, d2_mod_alt_23_25_June5,
  model_names = c("M1", "M2", "M3", "M4", "M5", "M6"), top_n = 5,
  title = "District 2 | Average Standardized Beta across Models", include_ecf = FALSE
)
res_beta$plot
```

### District 3

```{r blp3}
res_beta <- plot_avg_beta_lollipop_6model(
  d3_mod_time_23_25_June5, d3_mod_ln_price_no_ecf_23_25_June5,
  d3_mod_base_23_25_June5, d3_mod_base_inliers_23_25_June5,
  d3_mod_tasp_23_25_June5, d3_mod_alt_23_25_June5,
  model_names = c("M1", "M2", "M3", "M4", "M5", "M6"), top_n = 5,
  title = "District 3 | Average Standardized Beta across Models", include_ecf = FALSE
)
res_beta$plot
```

### District 4

```{r blp4}
res_beta <- plot_avg_beta_lollipop_6model(
  d4_mod_time_23_25_June2, d4_mod_ln_price_no_ecf_23_25_June2,
  d4_mod_base_23_25_June2, d4_mod_base_inliers_23_25_June2,
  d4_mod_tasp_23_25_June2, d4_mod_alt_23_25_June2,
  model_names = c("M1", "M2", "M3", "M4", "M5", "M6"), top_n = 5,
  title = "District 4 | Average Standardized Beta across Models", include_ecf = FALSE
)
res_beta$plot
```

### District 5

```{r blp5}
res_beta <- plot_avg_beta_lollipop_6model(
  d5_mod_time_23_25_June2, d5_mod_ln_price_no_ecf_23_25_June2,
  d5_mod_base_23_25_June2, d5_mod_base_inliers_23_25_June2,
  d5_mod_tasp_23_25_June2, d5_mod_alt_23_25_June2,
  model_names = c("M1", "M2", "M3", "M4", "M5", "M6"), top_n = 5,
  title = "District 5 | Average Standardized Beta across Models", include_ecf = FALSE
)
res_beta$plot
```

### District 6

```{r blp6}
res_beta <- plot_avg_beta_lollipop_6model(
  d6_mod_time_23_25_June2, d6_mod_ln_price_no_ecf_23_25_June2,
  d6_mod_base_23_25_June2, d6_mod_base_inliers_23_25_June2,
  d6_mod_tasp_23_25_June2, d6_mod_alt_23_25_June2,
  model_names = c("M1", "M2", "M3", "M4", "M5", "M6"), top_n = 5,
  title = "District 6 | Average Standardized Beta across Models", include_ecf = FALSE
)
res_beta$plot
```

### District 7

```{r blp7}
res_beta <- plot_avg_beta_lollipop_6model(
  d7_mod_time_23_25_June2, d7_mod_ln_price_no_ecf_23_25_June2,
  d7_mod_base_23_25_June2, d7_mod_base_inliers_23_25_June2,
  d7_mod_tasp_23_25_June2, d7_mod_alt_23_25_June2,
  model_names = c("M1", "M2", "M3", "M4", "M5", "M6"), top_n = 5,
  title = "District 7 | Average Standardized Beta across Models", include_ecf = FALSE
)
res_beta$plot
```




















