---
title: |
  | City Districts 1 - 7 Linear Regression Analysis 
  | Exploring Variable Impact on Sale Price | 2023 - 2025 Sales
author: |
  | Elliot Dean - Program Analyst IV  
  | Office of the Chief Financial Officer | Office of the Assessor
  | City of Detroit
date: "`r sub('^0', '', format(Sys.Date(), '%d %B %Y'))`"
geometry: margin = 1.5cm
output:
  pdf_document:
    latex_engine: xelatex
    toc: false
mainfont: Arial
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, error = FALSE, echo = FALSE)
```

```{r lib_load}
library(tidyverse)   # data wrangling & ggplot2    Comparison: R / SPSS
library(janitor)     # functions for examining and cleaning dirty data
library(kableExtra)  # simple table generator
library(here)        # Find files within a project
library(olsrr)       # package for p-value backward elimination
library(broom)       # summarizes key information about models
library(rlang)       # functions for Base Types and Core R and 'Tidyverse' Features
library(car)         # VIF, influence measures helpers
library(scales)      # provides the internal scaling infrastructure used by ggplot
```

```{r fun, results='hide'}
codf    <- readRDS(file = here("rmds/modfunctions", "codfunctions.rds"))
codform <- readRDS(file = here("rmds", "codformulas.rds"))
codf_env <- list2env(codf, parent = globalenv())
for (nm in names(codf)) {environment(codf[[nm]]) <- codf_env}
list2env(codf, envir = codf_env)
source(here("rmds/cod2325/districtruns/d2325f", "spss_cod_district-1_23-25f.R"), local = knitr::knit_global())
```

```{r setup1}
load(file = here("rmds/cod2325/districtruns/dmods", "spss_cod_district-1_23-25.RData"))
load(file = here("rmds/cod2325/districtruns/dmods", "spss_cod_district-2_23-25.RData"))
load(file = here("rmds/cod2325/districtruns/dmods", "spss_cod_district-3_23-25.RData"))
load(file = here("rmds/cod2325/districtruns/dmods", "spss_cod_district-4_23-25.RData"))
load(file = here("rmds/cod2325/districtruns/dmods", "spss_cod_district-5_23-25.RData"))
load(file = here("rmds/cod2325/districtruns/dmods", "spss_cod_district-6_23-25.RData"))
load(file = here("rmds/cod2325/districtruns/dmods", "spss_cod_district-7_23-25.RData"))
```

---

## Results Summary

The standardized beta metric is a unitless measure of the relationship between an independent
and dependent variable based on standard deviation (SD). The value of beta indicates that for every
increase in standard deviation of a predictor, the target will change by the equivalent amount.

The average values obtained by the backward linear regression runs in R suggest that the following
predictors have a higher likelihood of impacting sale price:

- Sale month (predominantly first half of the year)    | - Size
- Age                                                  | - Condition / Quality
- Exterior Siding Type                                 | - Restrooms
- Central Air Conditioning

In addition, a number of Economic Condition Factors were significantly correlated with sale price.
As these were District specific, they are reported after the shared predictors.\
* All figures represent models trained using ECFs (formulas 3 -> 6)

---

## Results Excluding Economic Condition Factors

```{r pplot1}
res_beta <- codf$plot_avg_beta_lollipop_7models(
  d1_mod_time_23_25_May12, d2_mod_time_23_25_May12, d3_mod_time_23_25_May12,
  d4_mod_time_23_25_May12, d5_mod_time_23_25_May12, d6_mod_time_23_25_May12,
  d7_mod_time_23_25_May12,
  model_names = c("D1", "D2", "D3", "D4", "D5", "D6", "D7"),
  top_n = 10,
  title = "LN_SPPSF Models - Standardized Beta Values"
)

res_beta$plot
```

```{r btab1}
res_extreme <- codf$summarize_extreme_betas_kable(
  models = list(d1_mod_time_23_25_May12, d2_mod_time_23_25_May12, d3_mod_time_23_25_May12,
  d4_mod_time_23_25_May12, d5_mod_time_23_25_May12, d6_mod_time_23_25_May12,
  d7_mod_time_23_25_May12),
  model_names = c("D1", "D2", "D3", "D4", "D5", "D6", "D7"),
  top_n = 3,
  bottom_n = 3,
  include_ecf = TRUE,
  caption = "Impact of Predictors on SPPSF"
) # Top and bottom standardized beta predictors across models

res_extreme$kable
```

---

```{r pplot2}
res_beta2 <- codf$plot_avg_beta_lollipop_7models(
  d1_mod_ln_price_no_ecf_23_25_May12, d2_mod_ln_price_no_ecf_23_25_May12, d3_mod_ln_price_no_ecf_23_25_May12,
  d4_mod_ln_price_no_ecf_23_25_May12, d5_mod_ln_price_no_ecf_23_25_May12, d6_mod_ln_price_no_ecf_23_25_May12,
  d7_mod_ln_price_no_ecf_23_25_May12,
  model_names = c("D1", "D2", "D3", "D4", "D5", "D6", "D7"),
  top_n = 10,
  title = "LN_PRICE (- ECF) Models | Standardized Beta Values"
)

res_beta2$plot
```

```{r btab2}
res_extreme_no_ecf <- codf$summarize_extreme_betas_kable(
  models = list(d1_mod_ln_price_no_ecf_23_25_May12, d2_mod_ln_price_no_ecf_23_25_May12,
                d3_mod_ln_price_no_ecf_23_25_May12, d4_mod_ln_price_no_ecf_23_25_May12,
                d5_mod_ln_price_no_ecf_23_25_May12, d6_mod_ln_price_no_ecf_23_25_May12,
                d7_mod_ln_price_no_ecf_23_25_May12),
  model_names = c("D1", "D2", "D3", "D4", "D5", "D6", "D7"),
  top_n = 3,
  bottom_n = 3,
  include_ecf = FALSE,
  caption = "Impact of Predictors on Saleprice (excluding ECFs)"
)

res_extreme_no_ecf$kable
```

---

```{r pplot3a, eval=FALSE}
res_beta3a <- codf$plot_avg_beta_lollipop_7models(
  d1_mod_base_23_25_May12, d2_mod_base_23_25_May12, d3_mod_base_23_25_May12,
  d4_mod_base_23_25_May12, d5_mod_base_23_25_May12, d6_mod_base_23_25_May12,
  d7_mod_base_23_25_May12,
  model_names = c("D1", "D2", "D3", "D4", "D5", "D6", "D7"),
  top_n = 10,
  title = "LN_PRICE Models - Standardized Beta Values"
)

res_beta3a$plot
```

```{r pplot3}
res_beta3 <- codf$plot_avg_beta_lollipop(
  models = list(d1_mod_base_23_25_May12, d2_mod_base_23_25_May12, d3_mod_base_23_25_May12,
                d4_mod_base_23_25_May12, d5_mod_base_23_25_May12, d6_mod_base_23_25_May12,
                d7_mod_base_23_25_May12),
  model_names = c("D1", "D2", "D3", "D4", "D5", "D6", "D7"),
  include_ecf = FALSE,
  top_n = 10,
  title = "LN_PRICE Models - Standardized Beta Values"
)

res_beta3$plot
```

```{r btab3}
res_extreme3 <- codf$summarize_extreme_betas_kable(
  models = list(d1_mod_base_23_25_May12, d2_mod_base_23_25_May12, d3_mod_base_23_25_May12,
                d4_mod_base_23_25_May12, d5_mod_base_23_25_May12, d6_mod_base_23_25_May12,
                d7_mod_base_23_25_May12),
  model_names = c("D1", "D2", "D3", "D4", "D5", "D6", "D7"),
  top_n = 3,
  bottom_n = 3,
  include_ecf = FALSE,
  caption = "Impact of Predictors on Saleprice"
)

res_extreme3$kable
```

---

```{r pplot4}
res_beta4 <- codf$plot_avg_beta_lollipop(
  models = list(d1_mod_base_inliers_23_25_May12, d2_mod_base_inliers_23_25_May12,
                d3_mod_base_inliers_23_25_May12, d4_mod_base_inliers_23_25_May12,
                d5_mod_base_inliers_23_25_May12, d6_mod_base_inliers_23_25_May12,
                d7_mod_base_inliers_23_25_May12),
  model_names = c("D1", "D2", "D3", "D4", "D5", "D6", "D7"),
  include_ecf = FALSE,
  top_n = 10,
  title = "LN_PRICE (Inliers) Models - Standardized Beta Values"
)

res_beta4$plot
```

```{r btab4}
res_extreme4 <- codf$summarize_extreme_betas_kable(
  models = list(d1_mod_base_inliers_23_25_May12, d2_mod_base_inliers_23_25_May12,
                d3_mod_base_inliers_23_25_May12, d4_mod_base_inliers_23_25_May12,
                d5_mod_base_inliers_23_25_May12, d6_mod_base_inliers_23_25_May12,
                d7_mod_base_inliers_23_25_May12),
  model_names = c("D1", "D2", "D3", "D4", "D5", "D6", "D7"),
  top_n = 3,
  bottom_n = 3,
  include_ecf = FALSE,
  caption = "Impact of Predictors on Saleprice"
)

res_extreme4$kable
```

---

```{r pplot5}
res_beta5 <- codf$plot_avg_beta_lollipop(
  models = list(d1_mod_tasp_23_25_May12, d2_mod_tasp_23_25_May12, d3_mod_tasp_23_25_May12,
                d4_mod_tasp_23_25_May12, d5_mod_tasp_23_25_May12, d6_mod_tasp_23_25_May12,
                d7_mod_tasp_23_25_May12),
  model_names = c("D1", "D2", "D3", "D4", "D5", "D6", "D7"),
  include_ecf = FALSE,
  top_n = 10,
  title = "LN_TASP Models - Standardized Beta Values"
)

res_beta5$plot
```

```{r btab5}
res_extreme5 <- codf$summarize_extreme_betas_kable(
  models = list(d1_mod_tasp_23_25_May12, d2_mod_tasp_23_25_May12, d3_mod_tasp_23_25_May12,
                d4_mod_tasp_23_25_May12, d5_mod_tasp_23_25_May12, d6_mod_tasp_23_25_May12,
                d7_mod_tasp_23_25_May12),
  model_names = c("D1", "D2", "D3", "D4", "D5", "D6", "D7"),
  top_n = 3,
  bottom_n = 3,
  include_ecf = FALSE,
  caption = "Impact of Predictors on TASP"
)

res_extreme5$kable
```

---

```{r pplot6}
res_beta6 <- codf$plot_avg_beta_lollipop(
  models = list(d1_mod_alt_23_25_May12, d2_mod_alt_23_25_May12, d3_mod_alt_23_25_May12,
                d4_mod_alt_23_25_May12, d5_mod_alt_23_25_May12, d6_mod_alt_23_25_May12,
                d7_mod_alt_23_25_May12),
  model_names = c("D1", "D2", "D3", "D4", "D5", "D6", "D7"),
  include_ecf = FALSE,
  top_n = 10,
  title = "Alternative LN_TASP Models - Standardized Beta Values"
)

res_beta6$plot
```

```{r btab6}
res_extreme6 <- codf$summarize_extreme_betas_kable(
  models = list(d1_mod_alt_23_25_May12, d2_mod_alt_23_25_May12, d3_mod_alt_23_25_May12,
                d4_mod_alt_23_25_May12, d5_mod_alt_23_25_May12, d6_mod_alt_23_25_May12,
                d7_mod_alt_23_25_May12),
  model_names = c("D1", "D2", "D3", "D4", "D5", "D6", "D7"),
  top_n = 3,
  bottom_n = 3,
  include_ecf = FALSE,
  caption = "Impact of Predictors on TASP"
)

res_extreme6$kable
```

---

## Results Including Economic Condition Factors

```{r pplot7}
res_beta7 <- codf$plot_avg_beta_lollipop(
  models = list(d1_mod_base_23_25_May12, d2_mod_base_23_25_May12, d3_mod_base_23_25_May12,
                d4_mod_base_23_25_May12, d5_mod_base_23_25_May12, d6_mod_base_23_25_May12,
                d7_mod_base_23_25_May12),
  model_names = c("D1", "D2", "D3", "D4", "D5", "D6", "D7"),
  include_ecf = TRUE,
  top_n = 10,
  title = "LN_PRICE Models - Standardized Beta Values"
)

res_beta7$plot
```

```{r btab7}
res_extreme7 <- codf$summarize_extreme_betas_kable(
  models = list(d1_mod_base_23_25_May12, d2_mod_base_23_25_May12, d3_mod_base_23_25_May12,
                d4_mod_base_23_25_May12, d5_mod_base_23_25_May12, d6_mod_base_23_25_May12,
                d7_mod_base_23_25_May12),
  model_names = c("D1", "D2", "D3", "D4", "D5", "D6", "D7"),
  top_n = 3,
  bottom_n = 3,
  include_ecf = TRUE,
  caption = "Impact of Predictors on Saleprice"
)

res_extreme7$kable
```

---

```{r pplot8}
res_beta8 <- codf$plot_avg_beta_lollipop(
  models = list(d1_mod_base_inliers_23_25_May12, d2_mod_base_inliers_23_25_May12,
                d3_mod_base_inliers_23_25_May12, d4_mod_base_inliers_23_25_May12,
                d5_mod_base_inliers_23_25_May12, d6_mod_base_inliers_23_25_May12,
                d7_mod_base_inliers_23_25_May12),
  model_names = c("D1", "D2", "D3", "D4", "D5", "D6", "D7"),
  include_ecf = TRUE,
  top_n = 10,
  title = "LN_PRICE (Inliers) Models - Standardized Beta Values"
)

res_beta8$plot
```

```{r btab8}
res_extreme8 <- codf$summarize_extreme_betas_kable(
  models = list(d1_mod_base_inliers_23_25_May12, d2_mod_base_inliers_23_25_May12,
                d3_mod_base_inliers_23_25_May12, d4_mod_base_inliers_23_25_May12,
                d5_mod_base_inliers_23_25_May12, d6_mod_base_inliers_23_25_May12,
                d7_mod_base_inliers_23_25_May12),
  model_names = c("D1", "D2", "D3", "D4", "D5", "D6", "D7"),
  top_n = 3,
  bottom_n = 3,
  include_ecf = TRUE,
  caption = "Impact of Predictors on Saleprice"
)

res_extreme8$kable
```

---

```{r pplot9}
res_beta9 <- codf$plot_avg_beta_lollipop(
  models = list(d1_mod_tasp_23_25_May12, d2_mod_tasp_23_25_May12, d3_mod_tasp_23_25_May12,
                d4_mod_tasp_23_25_May12, d5_mod_tasp_23_25_May12, d6_mod_tasp_23_25_May12,
                d7_mod_tasp_23_25_May12),
  model_names = c("D1", "D2", "D3", "D4", "D5", "D6", "D7"),
  include_ecf = TRUE,
  top_n = 10,
  title = "LN_TASP Models - Standardized Beta Values"
)

res_beta9$plot
```

```{r btab9}
res_extreme9 <- codf$summarize_extreme_betas_kable(
  models = list(d1_mod_tasp_23_25_May12, d2_mod_tasp_23_25_May12, d3_mod_tasp_23_25_May12,
                d4_mod_tasp_23_25_May12, d5_mod_tasp_23_25_May12, d6_mod_tasp_23_25_May12,
                d7_mod_tasp_23_25_May12),
  model_names = c("D1", "D2", "D3", "D4", "D5", "D6", "D7"),
  top_n = 3,
  bottom_n = 3,
  include_ecf = TRUE,
  caption = "Impact of Predictors on TASP"
)

res_extreme9$kable
```

---

```{r pplot10}
res_beta10 <- codf$plot_avg_beta_lollipop(
  models = list(d1_mod_alt_23_25_May12, d2_mod_alt_23_25_May12, d3_mod_alt_23_25_May12,
                d4_mod_alt_23_25_May12, d5_mod_alt_23_25_May12, d6_mod_alt_23_25_May12,
                d7_mod_alt_23_25_May12),
  model_names = c("D1", "D2", "D3", "D4", "D5", "D6", "D7"),
  include_ecf = TRUE,
  top_n = 10,
  title = "Alternative LN_TASP Models - Standardized Beta Values"
)

res_beta10$plot
```

```{r btab10}
res_extreme10 <- codf$summarize_extreme_betas_kable(
  models = list(d1_mod_alt_23_25_May12, d2_mod_alt_23_25_May12, d3_mod_alt_23_25_May12,
                d4_mod_alt_23_25_May12, d5_mod_alt_23_25_May12, d6_mod_alt_23_25_May12,
                d7_mod_alt_23_25_May12),
  model_names = c("D1", "D2", "D3", "D4", "D5", "D6", "D7"),
  top_n = 3,
  bottom_n = 3,
  include_ecf = TRUE,
  caption = "Impact of Predictors on TASP"
)

res_extreme10$kable
```













