---
title: "City of Detroit 2022 - 2025 Sales"
author: |
  | Elliot Dean - Program Analyst IV  
  | Office of the Chief Financial Officer | Office of the Assessor
  | City of Detroit
date: "`r sub('^0', '', format(Sys.Date(), '%d %B %Y'))`"
geometry: margin = 1.5cm
output:
  pdf_document:
    latex_engine: xelatex
    toc: true
mainfont: Arial
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, error = FALSE, echo = FALSE)
```

```{r lib_load}
library(tidyverse)   # data wrangling & ggplot2
library(janitor)     # functions for examining and cleaning dirty data
library(kableExtra)  # simple table generator
library(here)        # Find files within a project
library(olsrr)       # package for p-value backward elimination
library(broom)       # summarizes key information about models
library(rlang)       # functions for Base Types and Core R and 'Tidyverse' Features
library(car)         # VIF, influence measures helpers
library(scales)      # provides the internal scaling infrastructure used by ggplot
library(mgcv)        # Mixed GAM Computation Vehicle
library(fastDummies) # Fast Creation of Dummy (Binary) Columns & Rows from Categorical Variables
library(diffdf)      # tools for working with data frame difference
```

---

## Introduction

This report evaluates the metrics of a linear regression model created from citywide datasets of
sales data from the years 2022 - 2024 and 2023 - 2025. Overlapping sales IDs were removed to
include only unique sales, leaving a dataset containing single family residential sales from
2022 - 2025. After removing outliers in the initial data prep, a total of 19756 sales were used
for model development.

Custom functions were written to automate variable creation from initial datasets, calculate
assessing coefficients (PRD, COD, PRB), variance factors, and graphs.


```{r fun}
#load(file = here("rmds/modfunctions", "modfunctions11.RData"))
load(file = here("rmds/modfunctions", "modfunctions13.RData"))

codf    <- readRDS(file = here("rmds/modfunctions", "codfunctions.rds"))
codform <- readRDS(file = here("rmds", "codformulas.rds"))

codf_env <- list2env(codf, parent = globalenv())
for (nm in names(codf)) {environment(codf[[nm]]) <- codf_env}
list2env(codf, envir = codf_env)
source(here("rmds/cod2325/districtruns/d2325f", "spss_cod_district-7_23-25f.R"), local = knitr::knit_global())
```

Functions Used: modfunctions_notes.R + modelfunctions.R

```{r data_import}
load(file = here("data/rdata", "cod_2022-2025geo_23July.RData"))

df <- sales_df
```

```{r dataprep}
ecfs_col <- character(0)
if (isTRUE(make_ecf_dummies)) {
  if (!has_fastd) stop("make_ecf_dummies=TRUE requires fastDummies package.")
  if (!("ecf" %in% names(df))) stop("ecf column not found after rename; cannot dummy.")
  
  df <- fastDummies::dummy_cols(
    df, select_columns = "ecf", remove_first_dummy = remove_first_dummy,
    remove_selected_columns = !keep_ecf_original
  )
  ecfs_col <- grep("^ecf_", names(df), value = TRUE)
}
```

## Model I

```{r mod1}
# --- Model 1: LN_SPPSF with time splines (SPSS runs this to get time coefficients)
formula_time <- ln_sppsf ~ arterial +
  single_family1story + half_duplex1story + half_duplex2story +
  smallest + small + large + largest +
  two_plus_baths + powder_room +
  good + fair + poor + very_poor_unsound +
  qabove_avg + qbelow_avg +
  yb_pre_1929 + yb_1946_to_1960 + yb_post_1961 +
  crawl + slab +
  ln_garagespaces + fireplace +
  hot_water + elec_baseboard + wall + central_air +
  months_1to9 + months_10to16 + months_17to24 + near_downtown_1mi + near_downtown_5mi

mod_time <- backward_p(formula_time, data = df, p_out = 0.10, verbose = TRUE)
#summary(mod_time)

# --- Automatically pull time coefficients to compute RATEs (optional)
time_coefs <- coef(mod_time)[c("months_1to9", "months_10to16", "months_17to24")]
#time_coefs

rate1 <- unname(ifelse(is.na(time_coefs["months_1to9"]), 0, time_coefs["months_1to9"]))
rate2 <- unname(ifelse(is.na(time_coefs["months_10to16"]), 0, time_coefs["months_10to16"]))
rate3 <- unname(ifelse(is.na(time_coefs["months_17to24"]), 0, time_coefs["months_17to24"]))

get_b <- function(mod, nm) {
  cf <- stats::coef(mod)        # <- IMPORTANT: coef(mod), not coef[...]
  b  <- cf[nm]                  # returns NA if missing, never NULL
  base::unname(b)
}

b1 <- get_b(mod_time, "months_1to9")
b2 <- get_b(mod_time, "months_10to16")
b3 <- get_b(mod_time, "months_17to24")

# If you want to STOP when any term is missing:
missing <- c("months_1to9","months_10to16","months_17to24")[is.na(c(b1,b2,b3))]
if (length(missing) > 0) stop("Missing month term(s) in model: ", paste(missing, collapse=", "))

rate1 <- base::exp(b1) - 1
rate2 <- base::exp(b2) - 1
rate3 <- base::exp(b3) - 1

# PRICE_INDEX and TAF (time adjustment factor)
df <- df %>%
  mutate(
    price_index = (1 + rate1) ^ months_1to9 * (1 + rate2) ^ months_10to16 *
                  (1 + rate3) ^ months_17to24
  )

# Choose end_index = price_index at target month (SPSS: "last month")
# Here: by default, use the max MONTHS observed.
end_month <- max(df$months, na.rm = TRUE)
end_index <- df %>%  filter(months == end_month) %>%
  summarise(end_index = mean(price_index, na.rm = TRUE)) %>%  pull(end_index)

df <- df %>%
  mutate(
    taf   = end_index / price_index,    tasp  = saleprice * taf,
    tasppsf = tasp / sresb_floorarea,   ln_tasp = log(tasp)
  )
```

```{r mod1c, results='asis'}
model_metrics_A(mod_time, "LN\\_SPPSF model fit statistics")
coef_table(mod_time, "LN\\_SPPSF model coefficients (95% CI)")
```

---

## Model II

```{r mod2}
# --- Model 2: LN_PRICE without ECFs (SPSS "TEST MODEL WITHOUT ECFS")
formula_ln_price_no_ecf <- ln_price ~ ln_landratio + ln_base_bldgsize_ratio +
  arterial +
  single_family1story + half_duplex1story + half_duplex2story +
  smallest + small + large + largest +
  two_plus_baths + powder_room +
  good + fair + poor + very_poor_unsound +
  qabove_avg + qbelow_avg +
  yb_pre_1929 + yb_1946_to_1960 + yb_post_1961 +
  crawl + slab +
  ln_garagespaces + fireplace +
  hot_water + elec_baseboard + wall + central_air +
  months_1to9 + months_10to16 + months_17to24 + near_downtown_1mi + near_downtown_5mi

mod_ln_price_no_ecf <- backward_p(formula_ln_price_no_ecf, data = df, p_out = 0.10, verbose = TRUE)
# summary(mod_ln_price_no_ecf)
```

```{r modm2, results='asis'}
model_metrics_A(mod_ln_price_no_ecf, "LN\\_PRICE model fit statistics")
coef_table(mod_ln_price_no_ecf, "LN\\_PRICE model coefficients (95% CI)")
```

---

## Model III

```{r mod3}
# --- Model 3: BASE MODEL with ECFs (SPSS "BASE MODEL", dependent LN_PRICE)
# build RHS terms as character vector, then turn into a formula
rhs <- c(
  "ln_landratio", "ln_base_bldgsize_ratio",
  "arterial",
  "single_family1story", "half_duplex1story", "half_duplex2story",
  "smallest", "small", "large", "largest",
  "two_plus_baths", "powder_room",
  "good", "fair", "poor", "very_poor_unsound",
  "qabove_avg", "qbelow_avg",
  "yb_pre_1929", "yb_1946_to_1960", "yb_post_1961",
  "crawl", "slab",
  "ln_garagespaces", "fireplace",
  "hot_water", "elec_baseboard", "wall", "central_air",
  ecfs_col,  # <-- splices *all* secf_ dummy columns in correctly
  "months_1to9", "months_10to16", "months_17to24", "near_downtown_1mi", "near_downtown_5mi"
)

formula_base <- reformulate(rhs, response = "ln_price")

mod_base <- backward_p(formula_base, data = df, p_out = 0.10, verbose = TRUE)
#summary(mod_base)

# Save predictions and influence measures similar to SPSS /SAVE=
df <- df %>%
  mutate(pre_1 = predict(mod_base, newdata = df), esp1  = exp(pre_1), ratio1 = esp1 / saleprice)

# Assessing metrics output
rrmetrics <- ratio_metrics(df, esp1, saleprice)

knitr::kable(
  rrmetrics,  format   = "latex",  booktabs = TRUE,
  caption  = "LN\\_PRICE model (base) assessing coefficients"   # <-- no \\_
  # escape = TRUE is the default; keep it that way
) %>%  kable_styling(latex_options = c("hold_position", "striped"), font_size = 8)

# Ratio stats like SPSS RATIO STATS ... BY ...
ratio_ecf   <- ratio_by(df, esp1, saleprice, ecf)
ratio_style <- ratio_by(df, esp1, saleprice, sresb_style)
ratio_cond  <- ratio_by(df, esp1, saleprice, scond)
ratio_size  <- ratio_by(df, esp1, saleprice, size)
ratio_era   <- ratio_by(df, esp1, saleprice, erabuilt)
ratio_fire  <- ratio_by(df, esp1, saleprice, fireplace)
ratio_priceclass <- ratio_by(df, esp1, saleprice, priceclass)

# Within bands (10/20/50)
df <- df %>%
  mutate(
    within10_1 = if_else(ratio1 >= 0.9 & ratio1 <= 1.1, 1L, 0L),
    within20_1 = if_else(ratio1 >= 0.8 & ratio1 <= 1.2, 1L, 0L),
    within50_1 = if_else(ratio1 >= 0.5 & ratio1 <= 1.5, 1L, 0L)
  )

# 22) Outlier / influence removal (SPSS: COOK > 4/N, |SDRESID| > 2)
N_total <- nrow(df)
infl <- influence.measures(mod_base)
cook <- cooks.distance(mod_base)
stud <- rstudent(mod_base)

rows_used <- as.integer(rownames(model.frame(mod_base)))  # row numbers from the data used in lm()

df <- df %>%
  mutate(cook_1 = NA_real_,    stud_1  = NA_real_) |> 
  mutate(
    cook_1 = replace(cook_1, rows_used, cook),    stud_1 = replace(stud_1, rows_used, stud),
    in_inlier = if_else(cook_1 <= (4 / N_total) & abs(stud_1) <= 2, 1L, 0L)
    )

df_in <- df %>% filter(in_inlier == 1)

mod_base_inliers <- backward_p(formula_base, data = df_in, p_out = 0.10, verbose = TRUE)
#summary(mod_base_inliers)

df_in <- df_in %>%
  mutate(
    pre_2 = predict(mod_base_inliers, newdata = df_in),
    esp2  = exp(pre_2),    ratio2 = esp2 / saleprice,
    within10_2 = if_else(ratio2 >= 0.9 & ratio2 <= 1.1, 1L, 0L),
    within20_2 = if_else(ratio2 >= 0.8 & ratio2 <= 1.2, 1L, 0L),
    within50_2 = if_else(ratio2 >= 0.5 & ratio2 <= 1.5, 1L, 0L)
  )

# 23) Recalibrate time adjustments (SPSS shows an example with new RATEs)
# You can either:
#  - re-extract from a model that still has time splines, or  - set them manually.
# Here we demonstrate manual override:

time_coefs2 <- coef(mod_base)[c("months_1to9", "months_10to16", "months_17to24")]
#time_coefs2

# SPSS manually copies these:
# RATE1=-0.009842; RATE2=0.022764; RATE3=0
# If you'd rather hard-code like SPSS, set these explicitly.
RATE1_re <- unname(ifelse(is.na(time_coefs2["months_1to9"]), 0, time_coefs2["months_1to9"]))
RATE2_re <- unname(ifelse(is.na(time_coefs2["months_10to16"]), 0, time_coefs2["months_10to16"]))
RATE3_re <- unname(ifelse(is.na(time_coefs2["months_17to24"]), 0, time_coefs2["months_17to24"]))

df_in <- df_in %>%
  mutate(
    price_index_re = (1 + RATE1_re) ^ months_1to9 * (1 + RATE2_re) ^ months_10to16 *
                     (1 + RATE3_re) ^ months_17to24
  )

end_month_in <- max(df_in$months, na.rm = TRUE)
END_INDEX_re <- df_in %>%
  filter(months == end_month_in) %>%
  summarise(end_index = mean(price_index_re, na.rm = TRUE)) %>%  pull(end_index)

df_in <- df_in %>%
  mutate(
    taf_re  = END_INDEX_re / price_index_re,
    tasp_re = saleprice * taf_re,    ln_tasp_re = log(tasp_re)
  )

df_in1 <- df_in %>% select(saleid, pre_2, esp2, ratio2, within10_2, within20_2, within50_2,
                             price_index_re, taf_re, tasp_re, ln_tasp_re)

dfa <- df %>% left_join(df_in1, by = "saleid")
```

```{r modm3, results='asis'}
model_metrics_A(mod_base, "LN\\_PRICE (base) model fit statistics")
coef_table(mod_base, "LN\\_PRICE (+ECFs) model coefficients (95% CI)")
```

```{r m3vif, results='asis'}
safe_vif_kable(mod_base)
```

---

## Model IV

```{r mod4}
# 24) Model with recalibrated time adjustments (dependent LN_TASP, remove time vars)
# SPSS: LN_TASP with ECFs, no months splines
rhs2 <- c(
  "ln_landratio", "ln_base_bldgsize_ratio",
  "arterial",
  "single_family1story", "half_duplex1story", "half_duplex2story",
  "smallest", "small", "large", "largest",
  "two_plus_baths", "powder_room",
  "good", "fair", "poor", "very_poor_unsound",
  "qabove_avg", "qbelow_avg",
  "yb_pre_1929", "yb_1946_to_1960", "yb_post_1961",
  "crawl", "slab",
  "ln_garagespaces", "fireplace",
  "hot_water", "elec_baseboard", "wall", "central_air",
  ecfs_col, "near_downtown_1mi", "near_downtown_5mi"
)

formula_ln_tasp <- reformulate(rhs2, response = "ln_tasp_re")

mod_tasp <- backward_p(formula_ln_tasp, data = df_in, p_out = 0.10, verbose = TRUE)
#summary(mod_tasp)

df_in <- df_in %>%
  mutate(
    pre_3 = predict(mod_tasp, newdata = df_in), esp3  = exp(pre_3), ratio3 = esp3 / tasp_re
  )

rrmetrics <- ratio_metrics(df_in, esp3, saleprice)

knitr::kable(
  rrmetrics,  format   = "latex",  booktabs = TRUE,
  caption  = "LN\\_TASP model assessing coefficients" 
) %>%  kable_styling(latex_options = c("hold_position", "striped"), font_size = 8)

# 25) Alternative model (SPSS: LN_TASP with LN_LANDRATIO + LNSBLDSFINT + GARAGE + FIREPLACE, etc.)
rhs_alt <- c(
  "ln_landratio", "lnsbldsfint",
  "arterial",
  "single_family1story", "half_duplex1story", "half_duplex2story",
  "smallest", "small", "large", "largest",
  "two_plus_baths", "powder_room",
  "good", "fair", "poor", "very_poor_unsound",
  "qabove_avg", "qbelow_avg",
  "yb_pre_1929", "yb_1946_to_1960", "yb_post_1961",
  "crawl", "slab",
  "garage", "fireplace",
  "hot_water", "elec_baseboard", "wall", "central_air",
  ecfs_col, "near_downtown_1mi", "near_downtown_5mi"
)

formula_alt <- reformulate(rhs_alt, response = "ln_tasp_re")

mod_alt <- backward_p(formula_alt, data = df_in, p_out = 0.10, verbose = TRUE)
#summary(mod_alt)

df_in <- df_in %>%
  mutate(
    pre_4 = predict(mod_alt, newdata = df_in), esp4  = exp(pre_4), ratio4 = esp4 / tasp_re
  )

rrmetrics <- ratio_metrics(df_in, esp4, saleprice)

knitr::kable(
  rrmetrics,  format   = "latex",  booktabs = TRUE,
  caption  = "Alternative LN\\_TASP model assessing coefficients" 
) %>%  kable_styling(latex_options = c("hold_position", "striped"), font_size = 8)

df_in2 <- df_in %>% select(saleid, pre_3, esp3, ratio3, pre_4, esp4, ratio4)

dfa <- dfa %>% left_join(df_in2, by = "saleid")
```

```{r modm4, results='asis'}
model_metrics_A(mod_tasp, "LN\\_TASP model fit statistics")
coef_table(mod_tasp, "LN\\_TASP model coefficients (95% CI)")
```

```{r modm4a, results='asis'}
model_metrics_A(mod_alt, "Alternative LN\\_TASP model fit statistics")
coef_table(mod_alt, "Alternative LN\\_TASP model coefficients (95% CI)")
```

```{r m4vifs, results='asis'}
safe_vif_kable(mod_tasp)
safe_vif_kable(mod_alt)
```

```{r tplot}
ggplot(df_in, aes(x = tasp_re, y = esp3, colour = sresb_style)) +
  # points (open circles)
  geom_point(shape = 21, stroke = 0.8, fill = NA, alpha = 0.9, size = 2) +
  # 45° line
  geom_abline(intercept = 0, slope = 1) +
  # same scaling on both axes
  coord_equal() +
  # axis labels & legend title
  labs(x = "TASP",    y = "ESP3",    colour = "SRESB_STYLE") +
  # optional: colors roughly similar to your SPSS plot
  scale_colour_manual(
    values = c(
      "1/2 DUPLEX"    = "#66CCFF",  # light blue
      "SINGLE FAMILY" = "#CC3366"   # pink/magenta
    )
  ) +
  # theme similar to SPSS default
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.major = element_line(color = "grey80", linewidth = 0.3),
    panel.grid.minor = element_blank(),    legend.position  = "right",
    legend.title     = element_text(face = "bold")
  )
```

```{r datsav, eval=FALSE}
# dfc <- df
save(cod24, dfa, file = here("data/rdata", "cod_2022-2025.RData"))
```

```{r modsav}
cod_mod_time_22_25geo_July27 <- mod_time
cod_mod_ln_price_no_ecf_22_25geo_July27 <- mod_ln_price_no_ecf
cod_mod_base_22_25geo_July27 <- mod_base
cod_mod_base_inliers_22_25geo_July27 <- mod_base_inliers
cod_mod_tasp_22_25geo_July27 <- mod_tasp
cod_mod_alt_22_25geo_July27 <- mod_alt

save(cod_mod_time_22_25geo_July27, cod_mod_ln_price_no_ecf_22_25geo_July27,
     cod_mod_base_22_25geo_July27, cod_mod_base_inliers_22_25geo_July27,
     cod_mod_tasp_22_25geo_July27, cod_mod_alt_22_25geo_July27,
     file = here("data/rdata", "cod_2022-2025geo_July27.RData"))
```





















