---
title: "City of Detroit 2021 - 2026 Sales Workflow"
author: |
  | Elliot Dean - Program Analyst IV  
  | Office of the Chief Financial Officer | Office of the Assessor
  | City of Detroit
date: "`r sub('^0', '', format(Sys.Date(), '%d %B %Y'))`"
geometry: margin = 1.5cm
output:
  pdf_document:
    latex_engine: xelatex
    toc: true
mainfont: Arial
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, error = FALSE, echo = FALSE)
```

```{r lib_load}
library(tidyverse)    # data wrangling & ggplot2
library(janitor)      # functions for examining and cleaning dirty data
library(readxl)       # for reading Excel
library(kableExtra)   # simple table generator
library(here)         # Find files within a project
library(olsrr)        # package for p-value backward elimination
library(broom)        # summarizes key information about models
library(rlang)        # functions for Base Types and Core R and 'Tidyverse' Features
library(car)          # VIF, influence measures helpers
library(scales)       # provides the internal scaling infrastructure used by ggplot
library(mgcv)         # Mixed GAM Computation Vehicle
library(fastDummies)  # Fast Creation of Dummy (Binary) Columns & Rows from Categorical Variables
library(diffdf)       # tools for working with data frame difference
library(sf)           # used for handling and analyzing spatial data
library(httr2)        # make HTTP requests and handle responses
library(jsonlite)     # provides functions for parsing and generating JSON data
library(arcgislayers) # connect to ArcGIS Data Services
library(geosphere)    # Spherical trigonometry for geographic applications
library(osrm)         # Shortest Paths and Travel Time from OpenStreetMap
library(tigris)       # Load Census TIGER/Line Shapefiles
library(osmextract)   # Download and Import Open Street Map Data Extracts
```

---

## Introduction

This document contains the procedures implemented to import sales data into R, organize and clean
the dataset, create predictors for model training (including geospatial attributes), and save the
augmented dataset as an .RData file.

```{r fun}
load(file = here("rmds/modfunctions", "modfunctions13.RData"))

codf    <- readRDS(file = here("rmds/modfunctions", "codfunctions.rds"))
codform <- readRDS(file = here("rmds", "codformulas.rds"))

codf_env <- list2env(codf, parent = globalenv())
for (nm in names(codf)) {environment(codf[[nm]]) <- codf_env}
list2env(codf, envir = codf_env)
source(here("rmds/cod2325/districtruns/d2325f", "spss_cod_district-7_23-25f.R"), local = knitr::knit_global())
```

Functions Used: modfunctions_notes.R + modelfunctions.R

## Data Import

```{r data_import}
# ---- 1. Load data & normalize names ----
df <- read_csv(here("data/city", "SF-Sales-Apr2021-Mar2026-08182026.csv")) |>
  rename_with(tolower)    # clean_names()
```

## Prepare Dataset

```{r dataprep}
cod5     <- prep_cod_sales(df, ecf_col = "ecftbl", origin_month = as.Date("2021-04-01"))
df       <- cod5$data
ecfs_col <- cod5$ecfs_col
```

## Retrieve Parcel Coordinates

```{r addcoords}
df <- add_detroit_parcel_coords(
  df, "parcelno", x_col = "x_coord", y_col = "y_coord", batch_size = 100, quiet = FALSE
)
```

## Geospatial Attributes



### Distance to Downtown

```{r dist1}
df <- add_linear_distance(
  df = df,  lon_col = "x_coord",  lat_col = "y_coord",  target_lon = -83.0466,
  target_lat = 42.3317,  out_m_col = "dist_guardian_m",
  out_km_col = "dist_guardian_km", out_mi_col = "dist_guardian_mi"
)

df <- add_drive_distance_graphhopper(
  df = df,  lon_col = "x_coord",  lat_col = "y_coord",  target_lon = -83.0466,
  target_lat = 42.3317,  dist_col = "drive_guardian_km",  time_col = "drive_guardian_min",
  profile = "car", server_url = "http://localhost:8989"
)

df <- df %>%
  mutate(
    dist_guardian_log = log1p(dist_guardian_mi), drive_guardian_log = log1p(drive_guardian_min),
    near_downtown_1mi = as.integer(dist_guardian_mi <= 1),
    near_downtown_5mi = as.integer(dist_guardian_mi <= 5),
    near_downtown_10mi = as.integer(dist_guardian_mi <= 10),
  )
```










```{r datsav}
# dfc <- df
#save(sales_df, sales_df_in, file = here("data/rdata", "cod_2022-2025geofin.RData"))
save(df, file = here("data/rdata", "cod_2021-2026_19a.RData"))
```























