---
title: "Model Trial II"
author: "Elliot Dean"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  error   = FALSE
)
```

### Introduction

This is an initial test to integrate sales and property data from the .

------------------------------------------------------------------------

## Libraries
update.packages(ask = FALSE, checkBuilt = TRUE)

library(tidyverse) ;library(data.table) ; library(magrittr) ; library(janitor)  
*library(reticulate) ; *library(DBI) ; *library(odbc) ; *library(sf)
install.packages(c('tidyverse', 'caret', 'broom', 'car', 'lmtest', 'ggpubr', 'ggpmisc'))
install.packages(c('reticulate', 'DBI', 'odbc', 'sf', 'corrplot', 'httr', 'vip', 'here'))
install.packages(c("finetune", "DiceKriging", "rgenoud", 'leaps', 'nlstools', 'conflicted'))
install.packages("gtsummary", 'ggrepel', 'ggpp')

### Load Necessary Libraries

```{r lib_load}
library(tidyverse)   # data wrangling & ggplot2
library(magrittr)
library(janitor)
library(caret)       # train/test split, cross‐validation
library(broom)       # tidy model output
library(car)         # VIF, Durbin‐Watson
library(lmtest)      # Breusch–Pagan test
# library(data.table)
# library(reticulate)
```

### Importing and Organizing Data

Option 1, data.table, for larger sets, but I started with tidyverse below.

```{r, eval = FALSE}
# Import Parcel AND Sales Data using the fread function from the data.table package

parcelDatas <- fread('data/det_pars.csv')
salesDatas  <- fread('data/sales_det.csv')
```

START HERE:

```{r data_import}
# Import Parcel AND Sales Data using read_csv from readr package (Tidyverse)
# The clean_names method from janitor is used to standardized column names
parcels <- read_csv('data/ResiBldgData25.csv')  |> 
  clean_names()
sales   <- read_csv('data/SFResiSalesJUN2022-MAY2024_w-Exterior1.csv') |> 
  clean_names()
```

Input data (csv or otherwise) imported into RStudio can be saved as R objects for faster retrieval and streamlined workflow / organization:

```{r save_input, eval = FALSE}
# The save function is included in base R
# save(parcels, sales, file = 'data/spss_data.RData')
load('data/spss_data.RData')
```

The data contained in R objects can be retrieved with the load function:
load('data/spss_data.RData')

The data tables will be merged on Parcel Number (parcelno):

```{r data_merge}
df <- merge(sales, parcels, by = 'parcelno')
```

Remove duplicate entries (delete rows where parcel number appears twice or more):

```{r sel_distinct}
# Using distinct function, parcel column
# Created new dataframe (ddf) to maintain original
ddf <- df |> 
  distinct(parcelno, .keep_all = TRUE)
```

Make council district categorical for parameter testing:

```{r categorize}
ddf$cds <- as.character(ddf$s_council)
```

Drop duplicate columns (s_resb_garagearea, s_resb_floorarea, etc.):

```{r drop_dupes}
ddf <-
  subset(ddf, select = -c(s_resb_garagearea, s_resb_crawspace, s_resb_fullbaths,
                         s_resb_basementarea, s_resb_groundarea, s_resb_halfbaths,
                         s_resb_floorarea, s_resb_slabarea, s_resb_heat, s_cond,
                         s_resb_depr, s_aircond, s_esttcv, s_resb_bldgclass,
                         s_resb_yearbuilt, s_resb_style, s_resb_plusminus,
                         s_resb_occ))
```

Create a new column - Coordinates (lat_lon) - which merges X and Y coordinates.
Adding the Google Maps URL for API allows linking, Google Earth also added:

```{r coordinates}
ddf <- ddf |>  
  mutate(  # round to 6 decimals
    lat      = round(y, 6),
    lon      = round(x, 6),
    lat_lon  = paste(lat, lon, sep = ','), # combine into'42.334550,-83.043964'
    gmap_url = paste0(                     # build the Maps search URL
      'https://www.google.com/maps/search/?api=1&query=', lat_lon
    ),
    ge_url   = paste0(                     # build Google Earth search URL
      'https://earth.google.com/web/@', lat, ',', lon, ',1000a'
      # 1000 meters above ground (street-level), no tilt/bearing/etc.
    )
  # x = longitude, y = latitude; EPSG:4326 is standard WGS84
  # df_sf <- st_as_sf(df, coords = c("x", "y"), crs = 4326, remove = FALSE)
  )
```

Remove value, build, sqft, floor area outlier(s):

```{r data_filer}
ddf <- ddf |> 
  # drop_na() |> 
  filter(avwhensold < 1000000) |> 
  filter(yearbuilt > 1920 & yearbuilt < 2026) |> 
  filter(s_totalsqft < 100000 & s_totalsqft > 0) |>  
  filter(floorarea > 0) |> 
  # filter(cds != '22' & cds != '62') |> 
  filter(s_totalacres > 0)
```

Correctly format columns (date):

```{r data_form}
ddf$saledate <- as.Date(ddf$saledate, format = '%m/%d/%Y')
#ddf$sale_date.y <- sub('\\+00$', '', ddf$sale_date.y)
#ddf$sale_date.y <- as.Date(ymd_hms(ddf$sale_date.y, tz = 'UTC'))
```

Check dataframe data types using str function:

```{r check_df, eval = FALSE}
str(ddf)
```

As performed previously, merged datasets can also be saved as R objects for faster loading:

```{r save_dfs, eval = FALSE}
# The save function is included in base R
save(df, ddf, file = 'data/spss_frames.RData')
```

Add new columns, for data or organization:

```{r add_cols}
ddf <- ddf |> 
  mutate(numbaths = fullbaths + 0.5 * halfbaths)
```

Drop dupe columns + content

```{r drop_dupe_cols}
# Suppose df is your data.frame
# 1) Transpose to get columns as rows
mat <- t(as.matrix(df))

# 2) Identify duplicated rows (i.e., duplicate columns in original)
dups <- duplicated(mat)

# 3) Keep only non-duplicates
df_clean <- df[ , !dups]

# Drop dupe column names
# df <- df[ , !duplicated(names(df))]
```

LOAD DFS HERE:

```{r load_dfs, eval = FALSE}
load('data/spss_frames.RData')
```

Get some visualizations going

```{r vis1}
ggplot(
  data = ddf,
  mapping = aes(x = yearbuilt, y = avwhensold, color = cds)
) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    title = "Assessed Value by Build Year",
    subtitle = "...",
    x = "Year", y = "Value", color = "Ward"
  )
```

```{r vis2}
ggplot(
  data = ddf,
  mapping = aes(x = total_square_footage, y = taxable_value, color = zipc)
) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    title = "Taxable Value by Square Footage",
    subtitle = "...",
    x = "Area", y = "Value", color = "Council District"
  )
```

Manipulate data frames further as needed (remove outliers, etc.):

```{r data_filter2}
ddf2 <- ddf |> 
  # filter(sale_price.x < 10000000) 
  filter(year(sale_date.y) > 1950) |> 
  filter(sale_price.y < 20000000)  |> 
  filter(year(sale_date.x) > 1950) |> 
  filter(sale_price.x < 20000000)
```

```{r vis3}
ggplot(
  data = ddf2,
  mapping = aes(x = sale_price.x, y = sale_price.y, color = zipc)
) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    title = "Initial Sale Price versus Initial Sale Date",
    subtitle = "...",
    x = "Sale Date", y = "Sale Price", color = "Council District"
  )
```

Organize dataframe, look at stuff:

```{r plot_n_matrix}
# Scatterplots of price vs key predictors
ddf |>
  select(avwhensold, s_totalsqft, yearbuilt, numbaths, s_resb_nbed, cond) |>
  gather(key, val, -avwhensold) |>
  ggplot(aes(val, avwhensold)) +
    geom_point(alpha = .4) +
    facet_wrap(~key, scales = "free_x") +
    labs(x = NULL, y = "Assessed Value When Sold")

# Correlation matrix
ddf |> 
  select_if(is.numeric) |>
  cor(use = "pairwise.complete.obs") |>
  corrplot::corrplot(method = "ellipse")

```

Split into training & test sets:

```{r sets}
set.seed(123)
train_index <- createDataPartition(ddf$avwhensold, p = .7, list = FALSE)
train_data  <- ddf[train_index, ]
test_data   <- ddf[-train_index, ]
```

Fit a linear regression model:

```{r fit_lrm}
# Baseline: price ~ sqft_living + bedrooms + bathrooms
model1 <-
  lm(avwhensold ~ s_totalsqft + yearbuilt + numbaths + s_resb_nbed + cond,
     data = train_data)
summary(model1)
```

To add categorical variables or interactions:

```{r fit_lrm2}
model2 <- lm(price ~ sqft_living
                      + bedrooms
                      + bathrooms
                      + factor(grade)
                      + sqft_living:factor(zipcode),
             data = train_data)
summary(model2)
```

Examine model output with broom:

```{r broom}
# Coefficients & p‐values
tidy(model1)

# Overall model metrics (R², AIC, BIC)
glance(model1)
```

Check regression assumptions:

```{r rega}
# 1) Residual diagnostics
par(mfrow = c(2,2))
plot(model1)

# 2) Multicollinearity
vif(model1)   # VIF > 5–10 signals trouble

# 3) Heteroskedasticity
bptest(model1)   # Breusch–Pagan

# 4) Autocorrelation (if data are time-ordered)
durbinWatsonTest(model1)  
```

Evaluate on the test set:

```{r eval_ts}
pred <- predict(model1, newdata = test_data)
# Compute common metrics
postResample(pred = pred, obs = test_data$avwhensold)
#    RMSE    Rsquared      MAE 
#  123456   0.72         98765 
```

Cross-validation or regularization:

```{r cval_n_reg}
ctrl <- trainControl(method = "cv", number = 5)
lasso_fit <- train(avwhensold ~ ., data = train_data,
                   method = "glmnet",
                   trControl = ctrl,
                   tuneLength = 10)
lasso_fit
```

---

## Entire XGBoost Pipeline

```{r libsdatas}
# 1. Install and load necessary packages
# Ensure you have the required packages installed.
# install.packages(c("tidymodels", "AmesHousing", "dplyr", "ggplot2", "xgboost"))
library(tidymodels)
library(AmesHousing)
library(dplyr)
library(ggplot2)

# 2. Load and Inspect the Data
# We'll use the AmesHousing dataset, which is conveniently available in a package.
ames_raw <- make_ames()

# Get a glimpse of the data
glimpse(ames_raw)

# Summary of the target variable: Sale_Price
summary(ames_raw$Sale_Price)

# It's often better to predict the log of the sale price to handle skewed data
# and to penalize underestimates and overestimates more evenly.
ames_raw <- ames_raw %>%
  mutate(Sale_Price = log10(Sale_Price))
```

```{r splitfeats}
# 3. Initial Data Split
# We'll split the data into a training set and a testing set.
# The training set is used for model development and tuning.
# The testing set is held out until the very end to provide an unbiased
# estimate of the final model's performance.
set.seed(123) # for reproducibility
ames_split <- initial_split(ames_raw, prop = 0.80, strata = Sale_Price)

ames_train <- training(ames_split)
ames_test <- testing(ames_split)

# 4. Feature Engineering and Preprocessing with Recipes
# A 'recipe' is a series of steps for data preprocessing. This is one of the
# most powerful features of tidymodels. We will perform the following steps: 
# - Convert some integer variables that are actually categorical to factors.
# - Impute missing values.
# - Create dummy variables for categorical predictors.
# - Remove variables with zero variance (no information).
# - Normalize numeric predictors.

ames_recipe <- recipe(Sale_Price ~ ., data = ames_train) %>%
  # Impute missing numeric data with the median
  step_impute_median(all_numeric_predictors()) %>%
  # Impute missing categorical data with the most frequent value
  step_impute_mode(all_nominal_predictors()) %>%
  # Combine infrequent categorical levels into an "other" category. This helps
  # prevent issues with rare levels that might appear in one data split but not another.
  step_other(all_nominal_predictors(), threshold = 0.01) %>%
  # Create dummy variables for all categorical predictors
  step_dummy(all_nominal_predictors()) %>%
  # Remove predictors with no variance
  step_zv(all_predictors()) %>%
  # Normalize (center and scale) all numeric predictors
  step_normalize(all_numeric_predictors())

# You can "prep" and "bake" the recipe to see the results of the preprocessing
# This is an optional step for verification
prepped_recipe <- prep(ames_recipe, training = ames_train)
baked_data <- bake(prepped_recipe, new_data = ames_train)
glimpse(baked_data)
```

```{r modspecs}
# 5. Model Specification
# We'll use an XGBoost model, which is a powerful and popular algorithm for
# tabular data and often performs very well in these types of regression tasks.
# We will tune its hyperparameters.

xgb_spec <- boost_tree(
  trees = 1000,
  tree_depth = tune(),
  min_n = tune(),
  loss_reduction = tune(),
  sample_size = tune(),
  mtry = tune(),
  learn_rate = tune()
) %>%
  set_engine("xgboost") %>%
  set_mode("regression")

# 6. Create a Workflow
# A workflow bundles the recipe and the model specification together.
# This makes it much easier to manage the modeling process.

xgb_workflow <- workflow() %>%
  add_recipe(ames_recipe) %>%
  add_model(xgb_spec)

# 7. Hyperparameter Tuning
# We need to find the best combination of hyperparameters for our XGBoost model.
# We'll use cross-validation and a space-filling design to search the parameter space.
# Create cross-validation folds from the training data
set.seed(234)
ames_folds <- vfold_cv(ames_train, v = 5)

# Define the grid of hyperparameters to search
# We'll use a space-filling design to explore the parameter space efficiently.
set.seed(345)
xgb_grid <- grid_latin_hypercube(
  tree_depth(),
  min_n(),
  loss_reduction(),
  sample_size = sample_prop(),
  finalize(mtry(), ames_train),
  learn_rate(),
  size = 30 # Number of hyperparameter combinations to try
)

# Start the tuning process
# We'll use tune_grid() to apply our workflow to the cross-validation folds
# and test all the hyperparameter combinations.
# The `control_grid(save_pred = TRUE)` option saves the predictions for later inspection.
doParallel::registerDoParallel() # Use parallel processing to speed things up

set.seed(456)
xgb_tune_res <- tune_grid(
  xgb_workflow,
  resamples = ames_folds,
  grid = xgb_grid,
  control = control_grid(save_pred = TRUE)
)
```

```{r makedamod}
# 8. Analyze Tuning Results and Finalize the Model
# Let's see which hyperparameters performed the best.

# Show the best performing models based on root mean squared error (RMSE)
show_best(xgb_tune_res, metric = "rmse")

# We can also visualize the results
autoplot(xgb_tune_res)

# Select the best set of hyperparameters
best_xgb_params <- select_best(xgb_tune_res, metric = "rmse")

# Finalize the workflow by updating it with the best hyperparameters
final_xgb_workflow <- finalize_workflow(
  xgb_workflow,
  best_xgb_params
)

# 9. Train the Final Model and Evaluate on the Test Set
# Now, we'll train our final model on the *entire training set* and then
# evaluate its performance on the *held-out test set*. This gives us an
# unbiased
# assessment of how our model is likely to perform on new, unseen data.
# The `last_fit()` function does this for us: it fits the final model on the
# full training set and evaluates it on the test set.
final_fit <- last_fit(final_xgb_workflow, ames_split)

# Collect the metrics from the test set evaluation
test_set_metrics <- collect_metrics(final_fit)
print(test_set_metrics)

# Collect the predictions on the test set
test_set_predictions <- collect_predictions(final_fit)
print(test_set_predictions)

# Since we predicted the log10 of the price, let's transform it back
# to the original scale for interpretation.
test_set_predictions <- test_set_predictions %>%
  mutate(
    Sale_Price = 10^Sale_Price,
    .pred = 10^.pred
  )

# Visualize the final predictions vs actual values
ggplot(test_set_predictions, aes(x = Sale_Price, y = .pred)) +
  geom_point(alpha = 0.5) +
  geom_abline(color = "red", lty = 2) +
  labs(
    title = "Actual vs. Predicted Sale Prices on the Test Set",
    x = "Actual Sale Price",
    y = "Predicted Sale Price"
  ) +
  scale_x_continuous(labels = scales::dollar_format()) +
  scale_y_continuous(labels = scales::dollar_format()) +
  coord_obs_pred() +
  theme_minimal()

# You can now use the `final_fit` object to make predictions on new data.
# The workflow is stored inside:
final_model_object <- extract_workflow(final_fit)
# saveRDS(final_model_object, "my_avm_model.rds")
# To make predictions on new data:
# loaded_model <- readRDS("my_avm_model.rds")
# new_predictions <- predict(loaded_model, new_ames_data)
```

## Linear Regression vs XGBoost model comparison

```{r mod_compare, eval = FALSE}
# install.packages(c("caret", "leaps", "xgboost", "dplyr"))
library(caret)
library(leaps)

set.seed(42)

# ----------------------------
# 0) Data: set up your dataset
# ----------------------------
# df: a data.frame with outcome and predictors
# Example placeholders:
# df <- your_data_frame

y_col   <- "LN_PRICE"           # outcome
#num_cols <- c("ABV", "IBU", "Price")     # numeric predictors
#cat_cols <- c("Beer_Type", "Brewery")    # categorical predictors

# Keep only used columns
df <- df[, c(y_col, 'CRAWL', 'SLAB', 'LN_LANDRATIO', 'ARTERIAL',
                   'SINGLE_FAMILY1STORY', 'SINGLE_FAMILY2STORY', 'HALF_DUPLEX1STORY', 'HALF_DUPLEX2STORY',
                   'SMALLEST', 'SMALL', 'LARGE', 'LARGEST', 'AVG',
                   'TWO_PLUS_BATHS', 'POWDER_ROOM', 'LN_GARAGESPACES', 'FIREPLACE',
                   'EXCELLENT', 'VERY_GOOD', 'GOOD', 'AVERAGE', 'FAIR', 'POOR', 'VERY_POOR_UNSOUND',
                   'QABOVE_AVG', 'QBELOW_AVG', 'Q_AVG',
                   'YB_PRE_1929', 'YB_1930_TO_1945', 'YB_1946_TO_1960', 'YB_POST_1961',
                   'HOT_WATER', 'ELEC_BASEBOARD', 'ELEC_WALL', 'WALL', 'CENTRAL_AIR', 'ELEC_RADIANT',
                   'RADIANT_FLOOR', 'SPACE_HEATER',
                   'MONTHS_1TO9', 'MONTHS_10TO16', 'MONTHS_17TO24')] # num_cols', 'cat_cols

# Make sure outcome is numeric and cats are factors
df[[y_col]] <- as.numeric(df[[y_col]])
#for (cc in cat_cols) df[[cc]] <- as.factor(df[[cc]])

# caret formula interface (caret will handle factor dummies)
form <- as.formula(paste(y_col, "~ ."))

# ------------------------------------------
# 1) Define consistent outer CV (5-fold CV)
# ------------------------------------------
K_OUTER <- 5
outer_folds <- createFolds(df[[y_col]], k = K_OUTER, returnTrain = TRUE)

# Storage for outer-fold results
res_outer <- data.frame(
  fold        = seq_len(K_OUTER),
  RMSE_linear = NA_real_, MAE_linear = NA_real_, R2_linear = NA_real_,
  RMSE_xgb    = NA_real_, MAE_xgb    = NA_real_, R2_xgb    = NA_real_
)

# ------------------------------------------
# 2) Inner CV and model specs
# ------------------------------------------
# Inner CV (for tuning / feature selection)
inner_ctrl <- trainControl(
  method      = "cv",
  number      = 3,
  verboseIter = FALSE
)

# Preprocess choices (center/scale; imputation optional if needed)
# If you have missing values, consider preProcess = c("medianImpute","center","scale")
pp <- c("center", "scale")

# ---- Model A: Backward linear regression (leaps::regsubsets) via caret ----
# 'nvmax' is the number of predictors in the final subset (after dummying).
# You can set a sequence up to, say, 40; caret will choose the best via inner CV.
lin_grid <- expand.grid(nvmax = 1:20)  # adjust upper bound to your feature count

# ---- Model B: XGBoost (xgbTree) ----
xgb_grid <- expand.grid(
  nrounds          = c(300, 600, 900),
  max_depth        = c(3, 5, 7),
  eta              = c(0.03, 0.1),
  gamma            = 0,                # keep simple first pass
  colsample_bytree = c(0.7, 1.0),
  min_child_weight = c(1, 5),
  subsample        = c(0.7, 1.0)
)

# ------------------------------------------
# 3) Nested CV loop
# ------------------------------------------
for (i in seq_len(K_OUTER)) {
  cat(sprintf("\n=== OUTER FOLD %d/%d ===\n", i, K_OUTER))
  tr_idx <- outer_folds[[i]]
  te_idx <- setdiff(seq_len(nrow(df)), tr_idx)

  train_dat <- df[tr_idx, , drop = FALSE]
  test_dat  <- df[te_idx, , drop = FALSE]

  # -- Model A: backward selection (inner tuning) --
  lin_fit <- train(
    form, data = train_dat,
    method     = "leapBackward",  # requires the 'leaps' package
    tuneGrid   = lin_grid,
    trControl  = inner_ctrl,
    preProcess = pp,
    metric     = "RMSE"
  )

  # Predict on outer test
  pred_lin <- predict(lin_fit, newdata = test_dat)
  # Evaluate
  lin_post <- postResample(pred_lin, test_dat[[y_col]])  # RMSE, R2
  MAE_lin  <- mean(abs(pred_lin - test_dat[[y_col]]))

  # -- Model B: XGBoost (inner tuning) --
  xgb_fit <- train(
    form, data = train_dat,
    method     = "xgbTree",
    tuneGrid   = xgb_grid,
    trControl  = inner_ctrl,
    preProcess = pp,
    metric     = "RMSE",
    verbose    = FALSE
  )

  pred_xgb <- predict(xgb_fit, newdata = test_dat)
  xgb_post <- postResample(pred_xgb, test_dat[[y_col]])
  MAE_xgb  <- mean(abs(pred_xgb - test_dat[[y_col]]))

  # Save outer-fold metrics
  res_outer$RMSE_linear[i] <- lin_post["RMSE"]
  res_outer$R2_linear[i]   <- lin_post["Rsquared"]
  res_outer$MAE_linear[i]  <- MAE_lin

  res_outer$RMSE_xgb[i]    <- xgb_post["RMSE"]
  res_outer$R2_xgb[i]      <- xgb_post["Rsquared"]
  res_outer$MAE_xgb[i]     <- MAE_xgb
}

# ------------------------------------------
# 4) Summary + paired differences
# ------------------------------------------
summ <- res_outer |>
  summarize(
    RMSE_linear = mean(RMSE_linear), RMSE_linear_sd = sd(RMSE_linear),
    MAE_linear  = mean(MAE_linear),  MAE_linear_sd  = sd(MAE_linear),
    R2_linear   = mean(R2_linear),   R2_linear_sd   = sd(R2_linear),
    RMSE_xgb    = mean(RMSE_xgb),    RMSE_xgb_sd    = sd(RMSE_xgb),
    MAE_xgb     = mean(MAE_xgb),     MAE_xgb_sd     = sd(MAE_xgb),
    R2_xgb      = mean(R2_xgb),      R2_xgb_sd      = sd(R2_xgb)
  )

print(summ)

# Paired fold-wise differences (XGB - Linear); positive RMSE diff means XGB worse
rmse_diff <- res_outer$RMSE_xgb - res_outer$RMSE_linear
mae_diff  <- res_outer$MAE_xgb  - res_outer$MAE_linear
r2_diff   <- res_outer$R2_xgb   - res_outer$R2_linear

cat("\nPaired differences across identical outer folds:\n")
cat(sprintf("  RMSE diff (XGB - Linear): mean=%.4f, sd=%.4f\n",
            mean(rmse_diff), sd(rmse_diff)))
cat(sprintf("  MAE  diff (XGB - Linear): mean=%.4f, sd=%.4f\n",
            mean(mae_diff),  sd(mae_diff)))
cat(sprintf("  R2   diff (XGB - Linear): mean=%.4f, sd=%.4f\n",
            mean(r2_diff),   sd(r2_diff)))

# Optional: simple paired t-tests on fold-wise differences
cat("\nPaired t-tests on outer-fold differences:\n")
print(t.test(rmse_diff, mu = 0, paired = FALSE))  # small K; interpret cautiously
print(t.test(mae_diff,  mu = 0, paired = FALSE))
print(t.test(r2_diff,   mu = 0, paired = FALSE))

```

## ChatGPT 5 Comparison Script

```{r chat_compare}
# ============================================================
# Compare Linear Regression vs XGBoost (tidymodels framework)
# ============================================================

set.seed(123)

# ---- USER INPUTS -------------------------------------------
df         <- df              # e.g., df <- mydata
target     <- "LN_PRICE"                       # outcome column name (string)
predictors <- c('CRAWL', 'SLAB', 'LN_LANDRATIO', 'ARTERIAL',
                   'SINGLE_FAMILY1STORY', 'SINGLE_FAMILY2STORY', 'HALF_DUPLEX1STORY', 'HALF_DUPLEX2STORY',
                   'SMALLEST', 'SMALL', 'LARGE', 'LARGEST', 'AVG',
                   'TWO_PLUS_BATHS', 'POWDER_ROOM', 'LN_GARAGESPACES', 'FIREPLACE',
                   'EXCELLENT', 'VERY_GOOD', 'GOOD', 'AVERAGE', 'FAIR', 'POOR', 'VERY_POOR_UNSOUND',
                   'QABOVE_AVG', 'QBELOW_AVG', 'Q_AVG',
                   'YB_PRE_1929', 'YB_1930_TO_1945', 'YB_1946_TO_1960', 'YB_POST_1961',
                   'HOT_WATER', 'ELEC_BASEBOARD', 'ELEC_WALL', 'WALL', 'CENTRAL_AIR', 'ELEC_RADIANT',
                   'RADIANT_FLOOR', 'SPACE_HEATER',
                   'MONTHS_1TO9', 'MONTHS_10TO16', 'MONTHS_17TO24')            # <- your chosen predictors (character vector)
# ------------------------------------------------------------

# (1) Keep only rows with a *finite* outcome (guards against -Inf/NaN after logs)
df <- df[is.finite(df[[target]]), , drop = FALSE]

# (2) Make a slim copy with only the target + chosen predictors
keep_cols <- intersect(colnames(df), c(target, predictors))
missing_in_df <- setdiff(c(target, predictors), keep_cols)
if (length(missing_in_df)) {
  stop("These columns are not in df: ", paste(missing_in_df, collapse = ", "))
}
df_small <- df[, keep_cols, drop = FALSE]

# ---- resampling ----
split  <- initial_split(df_small, prop = 0.80, strata = !!sym(target))
train  <- training(split)
test   <- testing(split)
folds  <- vfold_cv(train, v = 10, strata = !!sym(target))

# ---- recipe (shared by both models) ----
# - remove any remaining NAs
# - drop zero-variance columns (per-fold safe)
# - one-hot encode factors
# - normalize numeric predictors
rec <- recipe(as.formula(paste(target, "~ .")), data = train) %>%
  step_naomit(all_predictors(), all_outcomes()) %>%
  step_zv(all_predictors()) %>%
  step_dummy(all_nominal_predictors(), one_hot = TRUE) %>%
  step_normalize(all_numeric_predictors())

# ============================================================
# Model 1: Linear regression (no tuning)
# ============================================================
spec_lm <- linear_reg() %>%
  set_engine("lm")

wf_lm <- workflow() %>%
  add_model(spec_lm) %>%
  add_recipe(rec)

# Cross-validated performance
lm_cv <- fit_resamples(
  wf_lm,
  resamples = folds,
  metrics   = metric_set(rmse, rsq, mae),
  control   = control_resamples(save_pred = TRUE)
)

# ============================================================
# Model 2: XGBoost regression (tuned)
# ============================================================
spec_xgb <- boost_tree(
  mode           = "regression",
  trees          = tune(),      # nrounds
  tree_depth     = tune(),
  learn_rate     = tune(),
  loss_reduction = tune(),      # gamma
  min_n          = tune(),      # min_child_weight (mapped)
  mtry           = tune(),      # colsample_bytree
  sample_size    = tune()       # subsample
) %>%
  set_engine("xgboost")

wf_xgb <- workflow() %>%
  add_model(spec_xgb) %>%
  add_recipe(rec)

# Parameter grid (data-aware for mtry)
pset <- parameters(wf_xgb) %>%
  # sensible bounds; adjust if needed
  update(
    trees          = trees(c(200, 1500)),
    tree_depth     = tree_depth(c(2, 10)),
    learn_rate     = learn_rate(c(-4, -1)),   # on log10 scale; converted by dials
    loss_reduction = loss_reduction(c(-6, 3)),
    min_n          = min_n(c(1, 20)),
    mtry           = finalize(mtry(), train),
    sample_size    = sample_prop(c(0.5, 1.0))
  )

grid <- grid_latin_hypercube(pset, size = 30)

xgb_tuned <- tune_grid(
  wf_xgb,
  resamples = folds,
  grid      = grid,
  metrics   = metric_set(rmse, rsq, mae),
  control   = control_grid(save_pred = TRUE)
)

best_xgb     <- select_best(xgb_tuned, metric = "rmse")
xgb_final_wf <- finalize_workflow(wf_xgb, best_xgb)

# ============================================================
# Test-set evaluation (fair holdout)
# ============================================================

# Linear: fit to full training set, evaluate on test
lm_last   <- last_fit(wf_lm, split, metrics = metric_set(rmse, rsq, mae))
xgb_last  <- last_fit(xgb_final_wf, split, metrics = metric_set(rmse, rsq, mae))

# --- Cross-validated metrics (train folds) ---
cat("\n=== Cross-validated metrics (train folds) ===\n")
cv_lm_tbl  <- collect_metrics(lm_cv)
print(cv_lm_tbl)

cat("\n--- XGBoost tuned (CV) ---\n")
cv_xgb_tbl <- collect_metrics(xgb_tuned) %>%
  dplyr::filter(.config == best_xgb$.config)
print(cv_xgb_tbl)

# --- Test-set metrics (holdout) ---
cat("\n=== Test-set metrics ===\n")

cat("\nLinear model on test:\n")
lm_test_tbl  <- collect_metrics(lm_last)   # columns: .metric, .estimator, .estimate
print(lm_test_tbl)

cat("\nXGBoost (tuned) on test:\n")
xgb_test_tbl <- collect_metrics(xgb_last)  # same columns
print(xgb_test_tbl)

# Side-by-side comparison for the test set
test_compare <- dplyr::bind_rows(
  lm_test_tbl  %>% dplyr::mutate(model = "linear"),
  xgb_test_tbl %>% dplyr::mutate(model = "xgboost")
) %>%
  dplyr::select(model, .metric, .estimate) %>%
  tidyr::pivot_wider(names_from = model, values_from = .estimate)

cat("\n=== Test-set comparison (wide) ===\n")
print(test_compare)

# ============================================================
# Inspect final fits / coefficients (optional)
# ============================================================
# Linear coefficients
lm_fit_obj <- extract_fit_parsnip(lm_last$.workflow[[1]])$fit
cat("\nLinear coefficients (test fit):\n")
print(coef(lm_fit_obj))

# XGBoost importances
xgb_fit_obj <- extract_fit_parsnip(xgb_last$.workflow[[1]])$fit
if (inherits(xgb_fit_obj, "xgb.Booster")) {
  cat("\nXGBoost feature importance (Gain):\n")
  print(xgboost::xgb.importance(model = xgb_fit_obj)[1:10, ])
}

```




```{r}
# --- Cross-validated metrics (train folds) ---
cat("\n=== Cross-validated metrics (train folds) ===\n")
cv_lm_tbl  <- collect_metrics(lm_cv)
print(cv_lm_tbl)

cat("\n--- XGBoost tuned (CV) ---\n")
cv_xgb_tbl <- collect_metrics(xgb_tuned) %>%
  dplyr::filter(.config == best_xgb$.config)
print(cv_xgb_tbl)

# --- Test-set metrics (holdout) ---
cat("\n=== Test-set metrics ===\n")

cat("\nLinear model on test:\n")
lm_test_tbl  <- collect_metrics(lm_last)   # columns: .metric, .estimator, .estimate
print(lm_test_tbl)

cat("\nXGBoost (tuned) on test:\n")
xgb_test_tbl <- collect_metrics(xgb_last)  # same columns
print(xgb_test_tbl)

# Side-by-side comparison for the test set
test_compare <- dplyr::bind_rows(
  lm_test_tbl  %>% dplyr::mutate(model = "linear"),
  xgb_test_tbl %>% dplyr::mutate(model = "xgboost")
) %>%
  dplyr::select(model, .metric, .estimate) %>%
  tidyr::pivot_wider(names_from = model, values_from = .estimate)

cat("\n=== Test-set comparison (wide) ===\n")
print(test_compare)
```


















