---
title:  'Model Trial I'
author: 'Elliot Dean'
date:   '`r Sys.Date()`'
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  error   = FALSE
)
```

### Introduction

This is an initial test to integrate sales and property data from the open source repository of the Assessors Office. [Property Sales](https://data.detroitmi.gov/datasets/27dc11249f5c41c0873449984986ea2f_3/explore?location=42.352654%2C-83.099544%2C10.33) and [Parcel](https://data.detroitmi.gov/datasets/3c784c118e5c4083b37038e9b38573df_0/explore?location=42.352624%2C-83.099130%2C10.57) data were obtained from the City of Detroit Open Data Portal. These files are named sales_det.csv and det_pars.csv respectively.

------------------------------------------------------------------------

### Libraries

library(tidyverse)  
library(data.table)  
library(magrittr)  
library(janitor)  
*library(reticulate)
*library(DBI)
*library(odbc)
*library(sf)
install.packages(c('tidyverse', 'caret', 'broom', 'car', 'lmtest', 'ggpubr', 'ggpmisc'))
install.packages(c('reticulate', 'DBI', 'odbc', 'sf', 'corrplot', 'httr'))

#### Load Necessary Libraries

```{r lib_load}
library(tidyverse)   # data wrangling & ggplot2
library(magrittr)
library(janitor)
library(caret)       # train/test split, cross‐validation
library(broom)       # tidy model output
library(car)         # VIF, Durbin‐Watson
library(lmtest)      # Breusch–Pagan test
# library(data.table)
# library(reticulate)
```

#### Importing and Organizing Data

Option 1, data.table, for larger sets, but I started with tidyverse below.

```{r, eval = FALSE}
# Import Parcel AND Sales Data using the fread function from the data.table package

parcelDatas <- fread('data/det_pars.csv')
salesDatas  <- fread('data/sales_det.csv')
```

START HERE:

Importing and Organizing Data

```{r data_import}
# Import Parcel AND Sales Data using read_csv from readr package (Tidyverse)
# The clean_names method from janitor is used to standardized column names
parcels_24 <- read_csv('data/DETPAR24.CSV')  |> 
  clean_names()
sales_24   <- read_csv('data/DETSALE24.CSV') |> 
  clean_names()

# Import SPSS Data using read_sav from haven package
# Mutate creates new columns that are functions of existing variables
data <- read_sav('data/v2Hoods_GrandmontAreaNewCond_Data.sav') |>
  mutate(across(where(is.labelled), as_factor)) |>
  clean_names()
```

Input data (as dataframes) can be saved as R objects

```{r save_input, eval = FALSE}
# The save function is included in base R
save(parcels, sales, data, file = 'data/input_data.RData')
```

The data contained in R objects can be retrieved with the load function

```{r load_input, eval = FALSE}
# The load function is included in base R
load('data/input_data.RData')
```


The data tables will be merged on 'parcel_number':

```{r data_merge}
df <- merge(sales, parcels, by = 'parcel_number')
```

Drop duplicate columns + content

```{r drop_dupe_cols}
# 1) Transpose to get columns as rows
mat <- t(as.matrix(df))

# 2) Identify duplicated rows (i.e., duplicate columns in original)
dups <- duplicated(mat)

# 3) Keep only non-duplicates
df_clean <- df[ , !dups]
```

Remove duplicate entries (delete rows where address appears twice)

```{r sel_distinct}
# Using distinct function, address column
# Created new dataframe (ddf) to maintain original
ddf <- df_clean |> 
  distinct(address, .keep_all = TRUE)
```

Make zip code, council district, and ward categorical for parameter testing

```{r categorize}
ddf$zipc  <- as.character(ddf$zip_code)
ddf$cds   <- as.character(ddf$council_district)
ddf$wards <- as.character(ddf$ward)
```

Create a new column - Coordinates (lat_lon) - which merges X and Y coordinates.
Adding the Google Maps URL for API allows linking, Google Earth also added:

```{r coordinates}
ddf <- ddf |>  
  mutate(  # round to 6 decimals
    lat      = round(y, 6),
    lon      = round(x, 6),
    lat_lon  = paste(lat, lon, sep = ','), # combine into'42.334550,-83.043964'
    gmap_url = paste0(                     # build the Maps search URL
      'https://www.google.com/maps/search/?api=1&query=', lat_lon
    ),
    ge_url   = paste0(                     # build Google Earth search URL
      'https://earth.google.com/web/@', lat, ',', lon, ',1000a'
      # 1000 meters above ground (street-level), no tilt/bearing/etc.
    )
  # x = longitude, y = latitude; EPSG:4326 is standard WGS84
  # df_sf <- st_as_sf(df, coords = c("x", "y"), crs = 4326, remove = FALSE)
  )
```

Remove value, build, sqft, floor area outlier(s):

```{r data_filer}
ddf <- ddf |> 
  filter(assessed_value < 1000000) |> 
  filter(year_built > 1920 & year_built < 2026) |> 
  filter(total_square_footage < 100000 & total_square_footage > 0) |>  
  filter(total_floor_area > 0) |> 
  filter(cds != '22' & cds != '62') |> 
  filter(total_acreage > 0) |> 
  filter(! zipc %in% c('482153', '48362', '84227', 'NA'))
```

Correctly format columns (Ward, date, etc):

```{r data_form}
ddf <- ddf |> 
  mutate(ward = sub('^0+', '', ddf$ward))

ddf$sale_date.x <- as.Date(ddf$sale_date.x, format = '%m/%d/%Y')
ddf$sale_date.y <- sub('\\+00$', '', ddf$sale_date.y)
ddf$sale_date.y <- as.Date(ymd_hms(ddf$sale_date.y, tz = 'UTC'))
```

Check dataframe data types using str function:

```{r check_df, eval = FALSE}
str(ddf)
```

As performed previously, merged datasets can also be saved as R objects for faster loading:

```{r save_dfs, eval = FALSE}
# The save function is included in base R
save(df, ddf, file = 'data/data_frames2.RData')
```

LOAD DFS HERE:

```{r load_dfs, eval = FALSE}
load('data/data_frames2.RData')
```

Get some visualizations going

```{r vis1}
ggplot(
  data = ddf,
  mapping = aes(x = year_built, y = assessed_value, color = wards)
) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    title = "Assessed Value by Build Year",
    subtitle = "...",
    x = "Year", y = "Value", color = "Ward"
  )
```

```{r vis2}
ggplot(
  data = ddf,
  mapping = aes(x = total_square_footage, y = taxable_value, color = zipc)
) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    title = "Taxable Value by Square Footage",
    subtitle = "...",
    x = "Area", y = "Value", color = "Council District"
  )
```

Manipulate data frames further as needed (remove outliers, etc.):

```{r data_filter2}
ddf2 <- ddf |> 
  # filter(sale_price.x < 10000000)
  # 
  filter(year(sale_date.y) > 1950) |> 
  filter(sale_price.y < 20000000)  |> 
  filter(year(sale_date.x) > 1950) |> 
  filter(sale_price.x < 20000000)
```


```{r vis3}
ggplot(
  data = ddf2,
  mapping = aes(x = sale_price.x, y = sale_price.y, color = zipc)
) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    title = "Initial Sale Price versus Initial Sale Date",
    subtitle = "...",
    x = "Sale Date", y = "Sale Price", color = "Council District"
  )
```

```{r vis4}
ggplot(
  data = ddf2,
  mapping = aes(x = sale_date.y, y = sale_price.y, color = cds)
) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    title = "Final Sale Price versus Final Sale Date",
    subtitle = "...",
    x = "Sale Date", y = "Sale Price", color = "Council District"
  )
```


```{r vis5}
ggplot(
  data = ddf,
  mapping = aes(x = total_acreage, y = taxable_value)
) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    title = "Taxable Value by Acreage",
    subtitle = "...",
    x = "Acres", y = "Value"
  )
```

Import different data files:

```{r data_importing}
# Import Parcel AND Sales Data using read_csv from readr package (Tidyverse)
# The clean_names method from janitor is used to standardized column names
parcel <- read_csv('data/EXPONEP.CSV')  |> 
  clean_names()
sale   <- read_csv('data/EXPONES.CSV') |> 
  clean_names()
```

Save new data as .RData:

```{r save_inputs}
# The save function is included in base R
save(parcel, sale, file = 'data/input_datas.RData')
```

Organize dataframe, look at stuff:

```{r plot_n_matrix}
# Scatterplots of price vs key predictors
ddf |>
  select(assessed_value, total_floor_area, year_built) |>
  gather(key, val, -assessed_value) |>
  ggplot(aes(val, assessed_value)) +
    geom_point(alpha = .4) +
    facet_wrap(~key, scales = "free_x") +
    labs(x = NULL, y = "Assessed Value")

# Correlation matrix
ddf |> 
  select_if(is.numeric) |>
  cor(use = "pairwise.complete.obs") |>
  corrplot::corrplot(method = "ellipse")

```

Split into training & test sets:

```{r sets}
set.seed(123)
train_index <- createDataPartition(ddf$assessed_value, p = .7, list = FALSE)
train_data  <- ddf[train_index, ]
test_data   <- ddf[-train_index, ]

```


Fit a linear regression model:

```{r fit_lrm}
# Baseline: price ~ sqft_living + bedrooms + bathrooms
model1 <- lm(assessed_value ~ total_floor_area + year_built, data = train_data)
summary(model1)


```

To add categorical variables or interactions:
```{r fit_lrm2}
model2 <- lm(price ~ sqft_living
                      + bedrooms
                      + bathrooms
                      + factor(grade)
                      + sqft_living:factor(zipcode),
             data = train_data)
summary(model2)

```


Examine model output with broom:

```{r broom}
# Coefficients & p‐values
tidy(model1)

# Overall model metrics (R², AIC, BIC)
glance(model1)

```

Check regression assumptions:

```{r rega}
# 1) Residual diagnostics
par(mfrow = c(2,2))
plot(model1)

# 2) Multicollinearity
vif(model1)   # VIF > 5–10 signals trouble

# 3) Heteroskedasticity
bptest(model1)   # Breusch–Pagan

# 4) Autocorrelation (if data are time-ordered)
durbinWatsonTest(model1)  

```


Evaluate on the test set:

```{r eval_ts}
pred <- predict(model1, newdata = test_data)

# Compute common metrics
postResample(pred = pred, obs = test_data$assessed_value)
#    RMSE    Rsquared      MAE 
#  123456   0.72         98765 

```

Cross-validation or regularization:

```{r cval_n_reg}
ctrl <- trainControl(method = "cv", number = 5)
lasso_fit <- train(assessed_value ~ ., data = train_data,
                   method = "glmnet",
                   trControl = ctrl,
                   tuneLength = 10)
lasso_fit

```







