---
title: "Modeling Motor City"
subtitle: "Building an Automated Valuation Model for the City of Detroit"
author: |
  | Elliot Dean - Program Analyst IV  
  | Office of the Chief Financial Officer | Office of the Assessor
  | City of Detroit
date: "`r sub('^0', '', format(Sys.Date(), '%d %B %Y'))`"
geometry: margin = 1.5cm
output:
  pdf_document:
    latex_engine: xelatex
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, error = FALSE)
```

---

## Abstract

The assessment of real estate is a complex process. Decades of research and protocol development
have provided an enormous volume of resources available to improve the efficiency and accuracy
of property valuation. As real estate markets can vary considerably between cities, the methods
to determine value are not ubiquitous. While organizations such as the International Association
of Assessing Officers (IAAO) have endeavored to recommend guidelines and approaches to valuation,
the vast diversity of markets restricts implementation of a singular framework. If real estate is
to be assessed equitably, the unique characteristics of the environment it is located in must be
considered. Variables such as population demographics, average age of properties, market condition,
legal statutes, and natural disaster risk may contribute to value. However, the impact of certain
variables on assessed value are not uniform - 

---

This is an initial analysis of the sales data (northeastCustomSales.xlsx) provided by
Monica Shihadeh on 2 September 2025. The excel file was converted to csv format,
using the read_excel() and write_csv() functions. The libraries utilized for this
analysis are listed below. Data was processed using similar methodology carried out
in SPSS. The SPSS code (v1Hoods_NE-Area_NewCond_Syntax.sps) served as a template.
Each block (chunk) of this document contains the sequential steps of the procedure.

Block names and overall purpose are as follows:

I.    lib_load      - activate libraries to be used
II.   data_import   - load the sales csv file into RStudio as a dataframe
III.  data_org_1    - creation of main variables and removing outliers
IV.   mod_make_1    - linear regression model for the first target
V.    data_org_2    - additional variable creation
VI.   mod_1_stats   - evaluation of the first model
VII.  mod_make_2    - linear regression model for the second target
VIII. mod_2_stats   - evaluation of the second model
IX.   mod_1_summary - detailed metrics of the first model
X.    mod_2_summary - detailed metrics of the second model
XI.   mod_save & mod_load - save and load models as R objects

The ultimate goal of this project is to produce a linear model that can
accurately predict the value of a residence from a given set of variables.
This analysis is limited in scope to the northeast area of Detroit.
As neighborhoods in the city vary considerably, multiple models will
eventually be created to fine-tune performance.


---

## Load Necessary Libraries

```{r lib_load}
library(tidyverse)   # data wrangling & ggplot2
library(janitor)     # functions for examining and cleaning dirty data
library(lmtest)      # Breusch–Pagan test
library(here)        # Find files within a project
library(car)         # vif()
library(ggpubr)      # Automatically print the regression equation for ggplot
library(ggrepel)     # provides geoms for ggplot2 to repel overlapping text labels
library(ggpmisc)     # Miscellaneous Extensions to 'ggplot2'
library(ggpp)        # Grammar Extensions to 'ggplot2'
library(spgwr)       # Functions for computing geographically weighted regressions
library(tidymodels)  # collection of packages for modeling and machine learning
library(bundle)      # Serialize Model Objects with a Consistent Interface
library(psych)       # group statistics, including correlations and factor analysis
library(kableExtra)  # make attractive tables
library(broom)       # Convert Statistical Objects into Tidy tables
library(scales)      # provides the internal scaling infrastructure used by ggplot2
library(olsrr)       # package for p-value backward elimination
library(MASS)        # Functions & datasets for Venables & Ripley
library(rlang)       # Functions for Base Types and Core R and 'Tidyverse' Features
library(gt)          # Easily create presentation-ready display tables
library(gtsummary)   # create publication-ready analytical and summary tables
library(skimr)       # Compact and Flexible Summaries of Data
```

### Create Custom Functions (Code in Appendix)

```{r load-funs, include = FALSE, ref.label = "functions"}
```

---

## Importing and Organizing Data

```{r data_import}
df <- read_csv(here("data", "allResiArmsLengthSalesSince2004.csv")) |> 
  rename_with(toupper)
```

---

## Total Residential Sales by Year

```{r total_sales}
date_col  <- "SALEDATE"         # datetime column
price_col <- "SALEPRICE"        # numeric sale price (if you want dollar volume)

df_clean <- df |>
  mutate(
    sale_date = as_date(!!sym(date_col)),    # safely convert to Date
    year      = year(sale_date)
  ) |> filter(!is.na(year))

by_year_n <- df_clean %>%
  count(year, name = "n_sales")

ggplot(by_year_n, aes(x = factor(year), y = n_sales, fill = n_sales)) +
  geom_col(width = 0.9) +
  #geom_smooth(aes(group=1), se=FALSE, col='black', size=0.75) +
  geom_text(aes(label = comma(n_sales)), vjust = -0.8, size = 7, fontface = "bold") +
  #geom_label(aes(label = comma(n_sales)), vjust = -0.8, size = 3, fill="white") +
  #geom_label_repel(aes(label = comma(n_sales)), force = 0.01, color = "white",
  #                 fill = "black", force_pull = 2, label.size = 0.1) +
  scale_y_continuous(labels = scales::comma, limits = c(0, 11000),
                     expand = expansion(mult = c(0, 0.12))) +
  scale_fill_viridis_c(option = "C", guide = "none") +
  labs(
    title = "Residential Property Sales by Year",
    subtitle = "Citywide Transaction Counts",
    x = "Year", y = "Number of Sales"
  ) +
  theme_minimal(base_size = 20) +
  theme(
    panel.grid.major.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    text = element_text(face = "bold"),         # everything bold...
    axis.title = element_text(face = "bold"),  # ...except axes
    axis.text  = element_text(face = "bold"),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  ) 
  

```


```{r}

scale_factor <- 1e6  # divide totals by 10^7

by_year_amt <- df_clean %>%
  filter(!is.na(.data[[price_col]])) %>%
  group_by(year) %>%
  summarise(total_sales = sum(.data[[price_col]], na.rm = TRUE), .groups = "drop") %>%
  mutate(total_scaled = total_sales / scale_factor)

pad <- 0.10 * max(by_year_amt$total_scaled, na.rm = TRUE)

ggplot(by_year_amt, aes(x = factor(year), y = total_scaled, fill = total_scaled)) +
  geom_col(width = 0.9) +
  geom_text_repel(
    aes(label = number(total_scaled, accuracy = 0.1)),  # 2 decimals; adjust as you like
    direction = "y",
    nudge_y = pad * 0.6,
    min.segment.length = 0,
    segment.size = 0,
    box.padding = 0.08,
    point.padding = 0.08,
    force = 3,
    max.overlaps = Inf,
    seed = 42,
    size = 7,
    fontface = "bold"
  ) +
  scale_y_continuous(
    labels = number_format(accuracy = 0.1),
    limits = c(0, max(by_year_amt$total_scaled) + pad),
    expand = expansion(mult = c(0, 0.02))
  ) +
  scale_fill_viridis_c(option = "C", guide = "none") +
  labs(
    title = "Residential Property Sales by Year",
    subtitle = "Total Value of Citywide Transactions",  # "Totals divided by 10^7 (values shown are in $×10^7)",
    x = "Year", y = "Total Sales (Million USD)"
  ) +
  theme_minimal(base_size = 20) +
  theme(
    text = element_text(face = "bold"),       # all bold...
    axis.title = element_text(face = "bold"),# ...except axes
    axis.text  = element_text(face = "bold"),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    panel.grid.major.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  coord_cartesian(clip = "off")

```



```{r}
by_year_amt <- df_clean %>%
  filter(!is.na(.data[[price_col]])) %>%
  group_by(year) %>%
  summarise(total_sales = sum(.data[[price_col]], na.rm = TRUE), .groups = "drop")

pad <- 0.1 * max(by_year_amt$total_sales, na.rm = TRUE)  # a bit more headroom for repel

ggplot(by_year_amt, aes(x = factor(year), y = total_sales, fill = total_sales)) +
  geom_col(width = 0.8) +
  geom_text_repel(
    aes(label = lab_sci_clean(digits = 3)(total_sales)),
    direction = "y",
    nudge_y = pad * 0.6,      # higher above bars
    min.segment.length = 0,
    segment.size = 0,
    box.padding = 0.08,
    point.padding = 0.08,
    force = 3,
    max.overlaps = Inf,
    seed = 42,
    size = 3,
    fontface = "bold"         # bold labels
  ) +
  scale_y_continuous(
    labels = lab_sci_clean(digits = 3),     # axis in clean scientific
    limits = c(0, max(by_year_amt$total_sales) + pad),
    expand = expansion(mult = c(0, 0.02))
  ) +
  scale_fill_viridis_c(option = "C", guide = "none") +
  labs(
    title = "Residential Property Sales by Year",
    subtitle = "Total dollar volume",
    x = NULL, y = "Total sales (scientific)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    text = element_text(face = "bold"),
    axis.title = element_text(face = "plain"),
    axis.text  = element_text(face = "plain"),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    panel.grid.major.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  coord_cartesian(clip = "off")
```





```{r}
ggplot(by_year_amt, aes(x = factor(year), y = total_sales, fill = total_sales)) +
  geom_col(width = 0.8) +
  #geom_text(aes(label = dollar(total_sales)), vjust = -0.3, size = 3) +
  geom_text(aes(label=formatC(total_sales, format = "e", digits = 2, drop0trailing = TRUE)), vjust = -0.3, size = 3) +
  scale_y_continuous(labels = dollar, expand = expansion(mult = c(0, 0.08))) +
  scale_fill_viridis_c(option = "C", guide = "none") +
  labs(
    title = "Residential Property Sales by Year",
    subtitle = "Total Market Value",
    x = "Year", y = "Total Sales (USD)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.major.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    text = element_text(face = "bold"),         # everything bold...
    axis.title = element_text(face = "plain"),  # ...except axes
    axis.text  = element_text(face = "plain"),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  )

ggplot(by_year_amt, aes(x = factor(year), y = total_sales, fill = total_sales)) +
  geom_col(width = 0.8) +
  # repel labels vertically, with a small upward nudge to keep distance from bars
  geom_text_repel(
    aes(label = label_scientific(digits = 3)(total_sales)),
    direction = "y",
    nudge_y = pad * 0.25,
    min.segment.length = 0,
    segment.size = 0,           # no leader lines
    box.padding = 0.05,
    point.padding = 0.05,
    force = 2,
    max.overlaps = Inf,
    seed = 42,
    size = 3
  ) +
  scale_y_continuous(
    labels = label_scientific(digits = 3),
    limits = c(0, max(by_year_amt$total_sales) + pad),
    expand = expansion(mult = c(0, 0.02))
  ) +
  scale_fill_viridis_c(option = "C", guide = "none") +
  labs(
    title = "Residential Property Sales by Year",
    subtitle = "Total dollar volume",
    x = NULL, y = "Total sales (scientific)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    text = element_text(face = "bold"),
    axis.title = element_text(face = "plain"),
    axis.text  = element_text(face = "plain"),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    panel.grid.major.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  coord_cartesian(clip = "off")   # let repel move a bit above the panel if needed

```




---

## Custom Functions

```{r functions, eval = FALSE}
gt_highlight_diffs <- function(data, col_a, col_b, n = 10, highlight = "#fff3bf",
                               title = "Difference Table") {        # "#ffe6e6"                     
  ca <- ensym(col_a); cb <- ensym(col_b)
  d <- data |>
    slice_head(n = n) |>
    mutate(
      .diff_row =
        (is.na(!!ca) != is.na(!!cb)) |                              # one NA, one not
        (!is.na(!!ca) & !is.na(!!cb) & (!!ca != !!cb))              # both present, not equal
    )
  gt_tbl <- d |> gt() |>
    tab_style(    # highlight col_a cells where the row differs
      style = list(cell_fill(color = highlight), cell_text(weight = "bold")),
      locations = cells_body(columns = !!ca, rows = .diff_row)
    ) |>
    tab_style(    # highlight col_b cells where the row differs
      style = list(cell_fill(color = highlight), cell_text(weight = "bold")),
      locations = cells_body(columns = !!cb, rows = .diff_row)
    ) |> cols_hide(.diff_row) |> tab_header(title = title)
  gt_tbl
}

plot_lm_pvalues_two <- function(mod1, mod2, labels = c("Model 1", "Model 2"),
                                drop_intercept = TRUE, digits = 3, alpha_line = 0.05,
                                show_labels = TRUE) {
  acc <- 10^(-digits)
  eps <- 1e-16
  td1 <- broom::tidy(mod1) |> mutate(.model = labels[1])
  td2 <- broom::tidy(mod2) |> mutate(.model = labels[2])
  df <- bind_rows(td1, td2)

  if (drop_intercept) df <- df |> filter(term != "(Intercept)")
  df <- df |>    # clamp tiny p for log scale; build formatted label
    mutate(p_clamp = pmax(p.value, eps), label_fmt = scales::number(p.value, accuracy = acc))
  
  order_df <- df |> # order terms by lowest p across models so rows are comparable
    group_by(term) |>
    summarise(p_min = min(p_clamp, na.rm = TRUE), .groups = "drop")

  df <- df |> left_join(order_df, by = "term") |>
    mutate(term_ord = forcats::fct_reorder(term, p_min, .desc = TRUE)) # least sig at top
  
  pd <- position_dodge(dodge_width)    # dodge so two model points don’t overlap
  p <- ggplot(df, aes(x = p_clamp, y = term_ord, colour = .model)) +
    geom_point(size = 3, position = pd) +
    geom_vline(xintercept = alpha_line, linetype = "dashed", colour = "red") +
    scale_x_log10(name = "P-Value (log scale)",
                  expand = expansion(mult = c(0.02, sidebar_space_mult))) +
    labs(title = "Predictor p-values by model", y = "Predictor", colour = "Model") + theme_minimal()

  if (show_labels) {
    if (requireNamespace("ggrepel", quietly = TRUE)) {
      p <- p + ggrepel::geom_text_repel(
        aes(label = label_fmt), size = 2.5, position = pd, max.overlaps = 200,
        box.padding = 0.25, point.padding = 0.2, min.segment.length = 0
      )
    } else {
      p <- p + geom_text(
        aes(label = label_fmt), size = 2.5, vjust = -0.6, position = pd, check_overlap = TRUE
      )
    }
  }
  p
}

plot_lm_pvalues_three <- function(mod1, mod2, labels = c("Model 1", "Model 2"), title = "Model",
                                     drop_intercept = TRUE, digits = 3, alpha_line = 0.05,
                                     show_labels = TRUE, label_style = c("decimal","scientific"),
                                     format_axis_scientific = TRUE, dodge_width = 0.5) {
  label_style <- match.arg(label_style)
  acc <- 10^(-digits)
  eps <- 1e-16
  td1 <- broom::tidy(mod1) |> mutate(.model = labels[1])
  td2 <- broom::tidy(mod2) |> mutate(.model = labels[2])
  df  <- bind_rows(td1, td2)

  if (drop_intercept) df <- df |> filter(term != "(Intercept)")
  lbl_fun <- if (label_style == "scientific") {    # choose a label formatter
    scales::label_scientific(digits = digits)
  } else {function(x) scales::number(x, accuracy = acc)}

  df <- df |> mutate(p_clamp = pmax(p.value, eps), label_fmt = lbl_fun(p.value))
  order_df <- df |>    # order rows by smallest p across models
    group_by(term) |>
    summarise(p_min = min(p_clamp, na.rm = TRUE), .groups = "drop")

  df <- df |>
    left_join(order_df, by = "term") |>
    mutate(term_ord = forcats::fct_reorder(term, p_min, .desc = TRUE))

  pd <- position_dodge(width = dodge_width)
  p <- ggplot(df, aes(x = p_clamp, y = term_ord, colour = .model)) +
    geom_point(size = 3, position = pd) +
    geom_vline(xintercept = alpha_line, linetype = "dashed", colour = "red") +
    scale_x_log10(
      name = "P-Value (log scale)",
      labels = if (format_axis_scientific) scales::label_scientific() else waiver()
    ) + labs(title = title, y = "Predictor", colour = "Model") +
    theme_minimal() + coord_cartesian(clip = "off")

  if (show_labels) {
    if (requireNamespace("ggrepel", quietly = TRUE)) {
      p <- p + ggrepel::geom_text_repel(aes(label = label_fmt),
        size = 2.5, position = pd, max.overlaps = 200,
        box.padding = 0.15, point.padding = 0.1, min.segment.length = 0
      )
    } else {
      p <- p + geom_text(
        aes(label = label_fmt), size = 2.5, vjust = -0.6, position = pd, check_overlap = TRUE
      )
    }
  }
  p
}

modmetrics <- function(model, title) {
  s1 <- summary(model)
  mod_df <- 
    data.frame(
      n           = nobs(model),
      p           = length(coef(model)), # number of parameters (incl. intercept)
      R2          = s1$r.squared,
      Adj_R2      = s1$adj.r.squared,
      ResStandDev = s1$sigma,  # residual std. dev. (RMSE-ish, with d.f. correction)
      ResDegFree  = model$df.residual,         # residual degrees of freedom
      RMSE        = sqrt(mean(residuals(model)^2)),
      MAE         = mean(abs(residuals(model))),
      AIC         = AIC(model)
  )
  gt(mod_df) |> tab_header(title = title)
}

# helper: clean scientific labels ("1e+05" -> "1e5")
lab_sci_clean <- function(digits = 3) {
  f <- scales::label_scientific(digits = digits)
  function(x) {
    s <- f(x)
    gsub("e\\+?0*", "e", s)  # drop '+' and zero padding in exponent
  }
}

```









