---
title: "Residential Sales - Exploratory Data Analysis"
author: |
  | Elliot Dean - Program Analyst IV  
  | Office of the Chief Financial Officer | Office of the Assessor
  | City of Detroit
date: "`r sub('^0', '', format(Sys.Date(), '%d %B %Y'))`"
geometry: margin = 1.5cm
output:
  pdf_document:
    latex_engine: xelatex
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  error   = FALSE
)
```

---

## Introduction

This report will highlight the impact of the physical condition classification rating on the
accuracy of linear regression models to predict sales price.
The files northeastCustomSales.csv and Dan_NorthEast_Sales1.csv were used for the sales data
compared in this analysis. The custom sales file contains updated condition classifications
for properties, the other csv contains older classification ratings.
The dataframes and models loaded here for analysis were produced by their respective R Markdown
documents (NE_Custom_Sales.Rmd, NE_Sales_Dan.Rmd).

## Load Necessary Libraries

```{r lib_load}
library(tidyverse)   # data wrangling & ggplot2
library(janitor)     # functions for examining and cleaning dirty data
library(lmtest)      # Breusch–Pagan test
library(here)        # Find files within a project
library(car)         # vif()
library(ggpubr)      # Automatically print the regression equation for ggplot
library(ggrepel)     # provides geoms for ggplot2 to repel overlapping text labels
library(ggpmisc)     # Miscellaneous Extensions to 'ggplot2'
library(ggpp)        # Grammar Extensions to 'ggplot2'
library(spgwr)       # Functions for computing geographically weighted regressions
library(tidymodels)  # collection of packages for modeling and machine learning
library(bundle)      # Serialize Model Objects with a Consistent Interface
library(psych)       # group statistics, including correlations and factor analysis
library(kableExtra)  # make attractive tables
library(broom)       # Convert Statistical Objects into Tidy tables
library(scales)      # provides the internal scaling infrastructure used by ggplot2
library(olsrr)       # package for p-value backward elimination
library(MASS)        # Functions & datasets for Venables & Ripley
library(rlang)       # Functions for Base Types and Core R and 'Tidyverse' Features
library(gt)          # Easily create presentation-ready display tables
library(gtsummary)   # create publication-ready analytical and summary tables
library(skimr)       # Compact and Flexible Summaries of Data
```

### Create Custom Functions (Code in Appendix)

```{r load-funs, include = FALSE, ref.label = "functions"}
```

## Importing and Organizing Data / Regression Models

```{r data_import}
load(here("data/northeast", "ne_dan_custom_sales_dfs.RData"))

A <- ne_custom_df
B <- ne_dan_df

clns <- readRDS(here("models", "necs_model_1.rds"))
clnp <- readRDS(here("models", "necs_model_2.rds"))
dlns <- readRDS(here("models", "nedan_model_1.rds"))
dlnp <- readRDS(here("models", "nedan_model_2.rds"))
```

---

## Compare Condition Ratings

```{r rating_comparison}
want <- c('SALEID', 'SALEDATE', 'SALEPRICE', 'SCOND')
small_A <- A |> select(all_of(want))   # error if any are absent
small_B <- B |> select(any_of(want))   # silently drops absent cols

x1 <- small_A |>
  arrange(SALEID, desc(SALEDATE)) |>        # choose your rule
  distinct(SALEID, .keep_all = TRUE)

y1 <- small_B |> arrange(SALEID, desc(SALEDATE)) |> distinct(SALEID, .keep_all = TRUE)

joined <- x1 |> full_join(y1, by = "SALEID")
ckeep <- c('SALEDATE', 'SALEPRICE', 'CUSTOM_CONDITION', 'DAN_CONDITION')

condComp <- joined |> 
  mutate(
    SALEDATE         = SALEDATE.x,
    SALEPRICE        = SALEPRICE.x,
    CUSTOM_CONDITION = SCOND.x,
    DAN_CONDITION    = SCOND.y
  )
conds <- condComp[ckeep]
gt_highlight_diffs(conds, CUSTOM_CONDITION, DAN_CONDITION,
                   title = "Example Condition Values from Each Sales Dataset")
```

---

## Compare Predictor Significance Values

```{r plotp}
plot_lm_pvalues_three(clns, dlns,
  labels = c("Custom","Dan"),
  title = "Predictor p-values by Model, SPPSF",
  label_style = "scientific",
  digits = 2,
  format_axis_scientific = TRUE
)

plot_lm_pvalues_three(clnp, dlnp,
  labels = c("Custom","Dan"),
  title = "Predictor p-values by Model, LN_PRICE",
  label_style = "scientific",
  digits = 2,
  format_axis_scientific = TRUE
)
```

---

## Compare Model Metrics

```{r modmetrics}
modmetrics(clns, "Custom Model (SPPSF) Metrics")
modmetrics(dlns, "Dan Model (SPPSF) Metrics")
modmetrics(clnp, "Custom Model (LN_PRICE) Metrics")
modmetrics(dlnp, "Dan Model (LN_PRICE) Metrics")
```

## Conclusion

While a perfect side-by-side comparison between the 2 models is not possible due to the
backward regression process eliminating insignificant predictors
(along with different time variables), the striking variation in adjusted R squared values and
degree of significance of predictors clearly shows the importance of accurately characterizing
the condition of a property. The most powerful predictors of property value were the 
condition variables. This is clearly demonstrated by both models, but the Custom condition
adjustments greatly improve the significance of certain condition factors compared to the
original values.

Acknowledging that property condition is subjectively assessed, it is clear that changing how
those determinations are made can impact model accuracy enormously. 

## Custom Functions

```{r functions, eval = FALSE}
gt_highlight_diffs <- function(data, col_a, col_b, n = 10, highlight = "#fff3bf",
                               title = "Difference Table") {        # "#ffe6e6"                     
  ca <- ensym(col_a); cb <- ensym(col_b)
  d <- data |>
    slice_head(n = n) |>
    mutate(
      .diff_row =
        (is.na(!!ca) != is.na(!!cb)) |                              # one NA, one not
        (!is.na(!!ca) & !is.na(!!cb) & (!!ca != !!cb))              # both present, not equal
    )
  gt_tbl <- d |> gt() |>
    tab_style(    # highlight col_a cells where the row differs
      style = list(cell_fill(color = highlight), cell_text(weight = "bold")),
      locations = cells_body(columns = !!ca, rows = .diff_row)
    ) |>
    tab_style(    # highlight col_b cells where the row differs
      style = list(cell_fill(color = highlight), cell_text(weight = "bold")),
      locations = cells_body(columns = !!cb, rows = .diff_row)
    ) |> cols_hide(.diff_row) |> tab_header(title = title) |> cols_align(align = "center") |>
    opt_table_outline() # |> opt_table_lines()
  gt_tbl
}

plot_lm_pvalues_two <- function(mod1, mod2, labels = c("Model 1", "Model 2"),
                                drop_intercept = TRUE, digits = 3, alpha_line = 0.05,
                                show_labels = TRUE) {
  acc <- 10^(-digits)
  eps <- 1e-16
  td1 <- broom::tidy(mod1) |> mutate(.model = labels[1])
  td2 <- broom::tidy(mod2) |> mutate(.model = labels[2])
  df <- bind_rows(td1, td2)

  if (drop_intercept) df <- df |> filter(term != "(Intercept)")
  df <- df |>    # clamp tiny p for log scale; build formatted label
    mutate(p_clamp = pmax(p.value, eps), label_fmt = scales::number(p.value, accuracy = acc))
  
  order_df <- df |> # order terms by lowest p across models so rows are comparable
    group_by(term) |>
    summarise(p_min = min(p_clamp, na.rm = TRUE), .groups = "drop")

  df <- df |> left_join(order_df, by = "term") |>
    mutate(term_ord = forcats::fct_reorder(term, p_min, .desc = TRUE)) # least sig at top
  
  pd <- position_dodge(dodge_width)    # dodge so two model points don’t overlap
  p <- ggplot(df, aes(x = p_clamp, y = term_ord, colour = .model)) +
    geom_point(size = 3, position = pd) +
    geom_vline(xintercept = alpha_line, linetype = "dashed", colour = "red") +
    scale_x_log10(name = "P-Value (log scale)",
                  expand = expansion(mult = c(0.02, sidebar_space_mult))) +
    labs(title = "Predictor p-values by model", y = "Predictor", colour = "Model") + theme_minimal()

  if (show_labels) {
    if (requireNamespace("ggrepel", quietly = TRUE)) {
      p <- p + ggrepel::geom_text_repel(
        aes(label = label_fmt), size = 2.5, position = pd, max.overlaps = 200,
        box.padding = 0.25, point.padding = 0.2, min.segment.length = 0
      )
    } else {
      p <- p + geom_text(
        aes(label = label_fmt), size = 2.5, vjust = -0.6, position = pd, check_overlap = TRUE
      )
    }
  }
  p
}

plot_lm_pvalues_three <- function(mod1, mod2, labels = c("Model 1", "Model 2"), title = "Model",
                                     drop_intercept = TRUE, digits = 2, alpha_line = 0.05,
                                     show_labels = TRUE, label_style = c("decimal","scientific"),
                                     format_axis_scientific = TRUE, dodge_width = 1.5) {
  label_style <- match.arg(label_style)
  acc <- 10^(-digits)
  eps <- 1e-16
  td1 <- broom::tidy(mod1) |> mutate(.model = labels[1])
  td2 <- broom::tidy(mod2) |> mutate(.model = labels[2])
  df  <- bind_rows(td1, td2)

  if (drop_intercept) df <- df |> filter(term != "(Intercept)")
  lbl_fun <- if (label_style == "scientific") {    # choose a label formatter
    scales::label_scientific(digits = digits, drop0trailing = TRUE)
  } else {function(x) scales::number(x, accuracy = acc)}

  df <- df |> mutate(p_clamp = pmax(p.value, eps), label_fmt = lbl_fun(p.value))
  order_df <- df |>    # order rows by smallest p across models
    group_by(term) |>
    summarise(p_min = min(p_clamp, na.rm = TRUE), .groups = "drop")

  df <- df |>
    left_join(order_df, by = "term") |>
    mutate(term_ord = forcats::fct_reorder(term, p_min, .desc = TRUE))

  pd <- position_dodge(width = dodge_width)
  p <- ggplot(df, aes(x = p_clamp, y = term_ord, colour = .model)) +
    geom_point(size = 2.5, position = pd) +
    geom_vline(xintercept = alpha_line, linetype = "dashed", colour = "red") +
    scale_x_log10(
      name = "P-Value (log scale)",
      labels = if (format_axis_scientific) scales::label_scientific() else waiver()
    ) + labs(title = title, y = "Predictor", colour = "Model") +
    theme_minimal() + coord_cartesian(clip = "off")

  if (show_labels) {
    if (requireNamespace("ggrepel", quietly = TRUE)) {
      p <- p + ggrepel::geom_text_repel(aes(label = label_fmt),
        size = 3, position = pd, max.overlaps = 50,
        box.padding = 0.5, point.padding = 0.5, min.segment.length = 0
      )
    } else {
      p <- p + geom_text(
        aes(label = label_fmt), size = 3, vjust = -0.8, position = pd, check_overlap = TRUE
      )
    }
  }
  p
}

modmetrics <- function(model, title) {
  s1 <- summary(model)
  mod_df <- 
    data.frame(
      n           = nobs(model),
      p           = length(coef(model)), # number of parameters (incl. intercept)
      R2          = s1$r.squared,
      Adj_R2      = s1$adj.r.squared,
      ResStandDev = s1$sigma,  # residual std. dev. (RMSE-ish, with d.f. correction)
      ResDegFree  = model$df.residual,         # residual degrees of freedom
      RMSE        = sqrt(mean(residuals(model)^2)),
      MAE         = mean(abs(residuals(model))),
      AIC         = AIC(model)
  )
  gt(mod_df) |> tab_header(title = title)
}
```

