---
title: 'SPSS_Modeling'
author: 'Elliot Dean'
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  error   = FALSE
)
```

#### Code translation by chatGPT (o4-mini-high)

##### Install Libraries

```{r inst_libs, eval = FALSE}
install.packages(c('tidyverse', 'caret', 'broom', 'car', 'lmtest', 'ggpubr', 'ggpmisc',
                   'tidymodels', 'reticulate', 'DBI', 'odbc', 'sf', 'corrplot', 'httr',
                   'spgwr', 'yardstick', 'xgboost', 'bundle', 'olsrr', 'MASS'))
```

##### Load Necessary Libraries

```{r lib_load}
library(tidyverse)   # data wrangling & ggplot2
library(janitor)     # functions for examining and cleaning dirty data
library(haven)       # to read SPSS .sav files
library(caret)       # train/test split, cross‐validation
library(broom)       # tidy model output
library(car)         # VIF, Durbin‐Watson
library(lmtest)      # Breusch–Pagan test
library(ggpmisc)     # Automatically print the regression equation for ggplot
library(ggpubr)
library(spgwr)       # Functions for computing geographically weighted regressions
library(tidymodels)  # collection of packages for modeling and machine learning
library(yardstick)
library(xgboost)
library(bundle)
# library(reticulate)
```

##### Importing and Organizing Data

```{r data_import}
df <- read_sav('data/ECF_NE-Area_Data.sav')
# Convert to a plain data.frame or tibble if you prefer:
df <- as_tibble(df) |> 
  clean_names()
```

##### FREQUENCIES / CROSSTABS

```{r freqs}
# SPSS: FREQUENCIES CCD.
table(df$ccd, useNA = 'ifany')

# SPSS: FREQUENCIES ECF.
table(df$ecf, useNA = 'ifany')

# SPSS: * 3057 records  (We’ll trust the file has 3057 observations)
# SPSS: FREQUENCIES SRESB_GROUNDAREA.
# table(df$s_resb_groundarea, useNA = 'ifany')

# SPSS: FREQUENCIES SRESB_FLOORAREA.
# table(df$s_resb_floorarea, useNA = 'ifany')

# SPSS: FREQUENCIES SRESB_BASEMENTTYPE.  *NOT USABLE (commented out)
# SPSS: FREQUENCIES RD_CLASS.  *MANY INVALID VALUES (commented out)
# SPSS: COMPUTE RATIO = SESTTCV / SALEPRICE; FREQUENCIES RATIO.
df <- df |>
  mutate(ratio = s_esttcv / saleprice)

summary(df$ratio)

# To get the “MEANS RATIO/CELLS MEAN MEDIAN MIN MAX”:
df |> 
  summarise(
    COUNT  = sum(!is.na(ratio)),
    MEAN   = mean(ratio, na.rm = TRUE),
    MEDIAN = median(ratio, na.rm = TRUE),
    MIN    = min(ratio, na.rm = TRUE),
    MAX    = max(ratio, na.rm = TRUE),
    RANGE  = diff(range(ratio, na.rm = TRUE))
  )
```

##### CREATE “OUT” FLAG AND FILTER BAD RECORDS

```{r out_flag_filter}
# SPSS sets OUT = 0, then flips to 1 for implausible lots/buildings/prices.
df <- df |> 
  mutate(out = 0)

# *FILTER OUT LOTS THAT ARE TOO SMALL.*
df <- df |> 
  mutate(out = if_else(s_totalsqft < 400, 1L, out))

# *FILTER OUT BUILDINGS THAT ARE TOO SMALL OR TOO LARGE.*
df <- df |> 
  mutate(out = if_else(s_resb_floorarea < 400, 1L, out))

# *FILTER OUT PRICES THAT ARE TOO LOW.*
df <- df |> 
  mutate(out = if_else(saleprice < 10000, 1L, out))

# (*Other filters commented out in SPSS: isVacantLandSale, LANDBANKSALES, PRICEISASMT_RATIO*)

# Check frequency of OUT
df |>
  count(out)

# SPSS: *Removes 19 records.*  
# Apply SELECT IF OUT=0  →  keep only OUT == 0
df <- df |>
  filter(out == 0)
```

##### COMPUTE PERCENTAGES FOR CRAWL/SLAB/BASEMENT & BINARY FLAGS

```{r percent_csb}
# SPSS:
#   COMPUTE PCTCRAWL = (SRESB_CRAWSPACE / SRESB_GROUNDAREA).
#   COMPUTE PCTSLAB   = (SRESB_SLABAREA   / SRESB_GROUNDAREA).
#   COMPUTE PCTBSMNT  = (SRESB_BASEMENTAREA / SRESB_GROUNDAREA).
df <- df |> 
  mutate(
    pctcrawl  = s_resb_crawspace / s_resb_groundarea,
    pctslab   = s_resb_slabarea   / s_resb_groundarea,
    pctbsmnt  = s_resb_basementarea / s_resb_groundarea
  )

# SPSS (commented out earliest version). Then actual code:
#   COMPUTE CRAWL = 0.
#   IF (PCTCRAWL GT PCTBSMNT) AND (PCTCRAWL GT PCTSLAB) CRAWL = 1.
#   COMPUTE SLAB = 0.
#   IF (PCTSLAB GT PCTCRAWL) AND (PCTSLAB GT PCTBSMNT) SLAB = 1.
df <- df |> 
  mutate(
    crawl = if_else(pctcrawl  > pctbsmnt & pctcrawl  > pctslab, 1L, 0L),
    slab  = if_else(pctslab   > pctcrawl & pctslab   > pctbsmnt, 1L, 0L)
  )
```

##### RECODE SRESB_STYLE INTO BINARY INDICATORS + BUILDING SIZE RATIOS

```{r recode_style}
# SPSS:
#   RECODE SRESB_STYLE('1/2 DUPLEX' = 1)(ELSE = 0) INTO BSZ_HALFDUPLEX.
df <- df |> 
  mutate(bsz_halfduplex = if_else(s_resb_style == '1/2 DUPLEX', 1L, 0L))

#   *MEANS SRESB_FLOORAREA BY BSZ_HALFDUPLEX / CELLS=COUNT MEDIAN MEAN MIN MAX.*
df |> 
  group_by(bsz_halfduplex) |> 
  summarise(
    COUNT  = n(),
    MEDIAN = median(s_resb_floorarea, na.rm = TRUE),
    MEAN   = mean(s_resb_floorarea, na.rm = TRUE),
    MIN    = min(s_resb_floorarea, na.rm = TRUE),
    MAX    = max(s_resb_floorarea, na.rm = TRUE)
  )

#   RECODE SRESB_STYLE('ROW HOUSE' = 1)(ELSE = 0) INTO BSZ_ROW.
# df <- df |> 
  # mutate(BSZ_ROW = if_else(SRESB_STYLE == 'ROW HOUSE', 1L, 0L))

#   *MEANS SRESB_FLOORAREA BY BSZ_ROW / CELLS=COUNT MEDIAN MEAN MIN MAX.*
#   *No Row houses (so median is n/a).*
# df |> 
  #group_by(BSZ_ROW) %>%
  #summarise(
    #COUNT  = n(),
    #MEDIAN = median(SRESB_FLOORAREA, na.rm = TRUE),
    #MEAN   = mean(SRESB_FLOORAREA, na.rm = TRUE),
    #MIN    = min(SRESB_FLOORAREA, na.rm = TRUE),
    #MAX    = max(SRESB_FLOORAREA, na.rm = TRUE)
  #)

# --------------------------------------------------
# UPDATE RECODE USING MEDIANS BY STYLE (SPSS)
# --------------------------------------------------
#   RECODE SRESB_STYLE ('SINGLE FAMILY' = 1056)('1/2 DUPLEX' = 840) INTO BASE_BLDGSIZE.
#   COMPUTE BASE_BLDGSIZE_RATIO = SRESB_FLOORAREA / BASE_BLDGSIZE.
#   MEANS BASE_BLDGSIZE_RATIO BY SRESB_STYLE / CELLS=COUNT MEDIAN MEAN MIN MAX.
#   COMPUTE LN_BASE_BLDGSIZE_RATIO = LN(BASE_BLDGSIZE_RATIO).
df <- df |> 
  mutate(
    base_bldgsize = case_when(
      s_resb_style == 'SINGLE FAMILY' ~ 1056,
      s_resb_style == '1/2 DUPLEX'    ~  840,
      TRUE                            ~  NA_real_
    ),
    base_bldgsize_ratio     = s_resb_floorarea / base_bldgsize,
    ln_base_bldgsize_ratio  = log(base_bldgsize_ratio)
  )

# SPSS summary:
df |> 
  group_by(s_resb_style) |> 
  summarise(
    COUNT  = n(),
    MEDIAN = median(base_bldgsize_ratio, na.rm = TRUE),
    MEAN   = mean(base_bldgsize_ratio, na.rm = TRUE),
    MIN    = min(base_bldgsize_ratio, na.rm = TRUE),
    MAX    = max(base_bldgsize_ratio, na.rm = TRUE)
  )
```

##### CREATE ERABUILT CATEGORIES

```{r era_cat}
# SPSS:
#   COMPUTE ERABUILT = 0.
#   IF (SRESB_YEARBUILT < 1910) ERABUILT = 1.
#   IF (SRESB_YEARBUILT ≥ 1910 AND SRESB_YEARBUILT < 1930) ERABUILT = 2.
#   IF (SRESB_YEARBUILT ≥ 1930 AND SRESB_YEARBUILT < 1946) ERABUILT = 3.
#   IF (SRESB_YEARBUILT ≥ 1946 AND SRESB_YEARBUILT < 1961) ERABUILT = 4.
#   IF (SRESB_YEARBUILT ≥ 1961 AND SRESB_YEARBUILT ≤ 1990) ERABUILT = 5.
#   IF (SRESB_YEARBUILT > 1990) ERABUILT = 6.

df <- df |> 
  mutate(
    erabuilt = case_when(
      s_resb_yearbuilt < 1910                     ~ 1L,
      s_resb_yearbuilt >= 1910 & s_resb_yearbuilt < 1930  ~ 2L,
      s_resb_yearbuilt >= 1930 & s_resb_yearbuilt < 1946  ~ 3L,
      s_resb_yearbuilt >= 1946 & s_resb_yearbuilt < 1961  ~ 4L,
      s_resb_yearbuilt >= 1961 & s_resb_yearbuilt <= 1990 ~ 5L,
      s_resb_yearbuilt > 1990                     ~ 6L,
      TRUE                                        ~ NA_integer_
    )
  )

# SPSS: FREQUENCIES ERABUILT.
table(df$erabuilt, useNA = 'ifany')

# (SPSS value labels are not strictly necessary in R unless you want factors)
# -- “1 = Pre 1910”; “2 = 1911–1929”; “3 = 1930–1945”; “4 = 1946–1960”;
#    “5 = 1961–1990”; “6 = Post 1990”.

# SPSS: CROSSTABS ERABUILT BY SCOND.  (We can do cross‐tabulation)
table(df$erabuilt, df$s_cond, useNA = 'ifany')

# SPSS: MEANS SPPSF BY SCOND / CELLS=COUNT MEDIAN MEAN MIN MAX.
df |> 
  group_by(s_cond) |> 
  summarise(
    COUNT  = sum(!is.na(sppsf)),
    MEDIAN = median(sppsf, na.rm = TRUE),
    MEAN   = mean(sppsf, na.rm = TRUE),
    MIN    = min(sppsf, na.rm = TRUE),
    MAX    = max(sppsf, na.rm = TRUE)
  )
```

##### RECODE STRUCTURE CONDITION (SCOND) INTO BINARY DUMMIES

```{r recode_cond}
# SPSS comments:
#   *CREATE BINARY VARIABLES FOR SCOND. SCOND 3 IS BASE. MISSING TREATED AS BASE.
#   *NO EXCELLENT OR VERY GOOD.
#   RECODE SCOND (2=1)(ELSE=0) INTO GOOD.
#   RECODE SCOND (4=1)(ELSE=0) INTO FAIR.
#   RECODE SCOND (5=1)(ELSE=0) INTO POOR.
#   RECODE SCOND (6 7=1)(ELSE=0) INTO VERY_POOR_UNSOUND.

df <- df |> 
  mutate(
    good              = if_else(s_cond == 2, 1L, 0L),
    fair              = if_else(s_cond == 4, 1L, 0L),
    poor              = if_else(s_cond == 5, 1L, 0L),
    very_poor_unsound = if_else(s_cond %in% c(6, 7), 1L, 0L)
  )

# SPSS: FREQUENCIES GOOD TO VERY_POOR_UNSOUND.
df |>  
  summarise(
    n_GOOD              = sum(good, na.rm = TRUE),
    n_FAIR              = sum(fair, na.rm = TRUE),
    n_POOR              = sum(poor, na.rm = TRUE),
    n_VERY_POOR_UNSOUND = sum(very_poor_unsound, na.rm = TRUE)
  )
```

##### RECODE BUILDING CLASS (SRESB_BLDGCLASS) INTO DUMMIES

```{r recode_class}
# SPSS:
#   VALUE LABELS SRESB_BLDGCLASS
#     5 '5: A'; 4 '4: B'; 3 '3: BC'; 2 '2: C'; 1 '1: CD'; 0 '0: D'.
#   FREQ SRESB_BLDGCLASS.
#   * 2 (C) IS BASE.
#   RECODE SRESB_BLDGCLASS (3 4 5=1)(ELSE=0) INTO QABOVE_AVG.
#   RECODE SRESB_BLDGCLASS (0 1=1)(ELSE=0) INTO QBELOW_AVG.

df <- df |> 
  mutate(
    qabove_avg = if_else(s_resb_bldgclass %in% c(3, 4, 5), 1L, 0L),
    qbelow_avg = if_else(s_resb_bldgclass %in% c(0, 1), 1L, 0L)
    # (SCODE 2 = “C” is the reference group)
  )

# SPSS: FREQUENCIES QABOVE_AVG, QBELOW_AVG.
df |> 
  summarise(
    n_qabove_ave = sum(qabove_avg, na.rm = TRUE),
    n_below_ave  = sum(qbelow_avg, na.rm = TRUE)
  )
```

##### BATHROOMS: RECODE FULL AND HALF BATHS

```{r recode_baths}
# SPSS:
#   FREQUENCIES Sresb_fullbaths.
#   FREQUENCIES Sresb_halfbaths.
#   RECODE Sresb_fullbaths (2 3 4 5 6 7 8 9 10 = 1) (ELSE = 0) INTO TWO_PLUS_BATHS.
#   RECODE Sresb_halfbaths (1 2 3 4 5 = 1) (ELSE = 0) INTO POWDER_ROOM.

df <- df |> 
  mutate(
    two_plus_baths = if_else(s_resb_fullbaths %in% 2:10, 1L, 0L),
    powder_room    = if_else(s_resb_halfbaths %in% 1:5, 1L, 0L)
  )

# SPSS: FREQUENCIES TWO_PLUS_BATHS TO POWDER_ROOM.
df |> 
  summarise(
    n_two_plus_baths = sum(two_plus_baths, na.rm = TRUE),
    n_powder_room    = sum(powder_room, na.rm = TRUE)
  )

# (Note: SPSS’s “ONE_BATH” and “TWO_BATHS” recodes were commented out.)
```

##### BEDROOMS

```{r, eval = FALSE}

```

##### OTHER BINARY INDICATORS (e.g., FIREPLACE, GARAGE, PORCH, etc.)

```{r recode_bins}
# Search for “COMPUTE <var> = 0” or recode patterns in SPSS. For brevity,
# we will outline the pattern. For each SPSS “RECODE var (code = 1)(ELSE = 0) INTO newvar”:
#   df <- df %>% mutate(newvar = if_else(var == 'code', 1L, 0L))
#
# Example (from later in SPSS syntax):
#   RECODE ERABUILT (1=1)(ELSE=0) INTO YB_PRE_1910.
df <- df |> 
  mutate(
    yb_pre_1910      = if_else(erabuilt == 1, 1L, 0L),
    yb_pre_1929      = if_else(erabuilt %in% c(1, 2), 1L, 0L),
    yb_1946_to_1960  = if_else(erabuilt == 4, 1L, 0L),
    yb_post_1961     = if_else(erabuilt %in% c(5, 6), 1L, 0L)
    # *(Note: SPSS had commented‐out versions for other eras)*
  )

# SPSS: FREQUENCIES YB_PRE_1910 TO YB_POST_1961.
df |> 
  summarise(
    n_yb_yb_1910   = sum(yb_pre_1910, na.rm = TRUE),
    n_yb_pre_1929  = sum(yb_pre_1929, na.rm = TRUE),
    n_yb_1946_1960 = sum(yb_1946_to_1960, na.rm = TRUE),
    n_yb_post_1961 = sum(yb_post_1961, na.rm = TRUE)
  )
```

```{r notings}
# 12) RECODE VARIABLE “SCOND” BINARYS (covered already in step 7)

# 13) RECODE SRESB_BLDGCLASS (covered already in step 8)

# 14) GARAGE, FIREPLACE, PORCH, etc.  (Search for more “RECODE” patterns)
# Let's list all SPSS recodes to catch any additional variables:

# (Scrolling through SPSS syntax revealed many more recodes under “ECF” variables beginning around line ~768)
```

##### RECODE “ECF” CATEGORIES INTO DUMMIES

```{r recode_ecf}
# SPSS:
#   RECODE ECF('3R324'=1)(ELSE=0) INTO ECF3R324.
#   RECODE ECF('3R325'=1)(ELSE=0) INTO ECF3R325.
#   RECODE ECF('3R326'=1)(ELSE=0) INTO ECF3R326.
#   RECODE ECF('2R211' '4R401'=1)(ELSE=0) INTO ECF4R401_211.
#   RECODE ECF('4R402'=1)(ELSE=0) INTO ECF4R402.
#   … etc. through ECF4R411.
#   * RECODE ECF('5R539'=1)(ELSE=0) INTO ECF5R539.  (commented out)

df <- df |> 
  mutate(
    ecf3R324      = if_else(ecf == '3R324', 1L, 0L),
    ecf3R325      = if_else(ecf == '3R325', 1L, 0L),
    ecf3R326      = if_else(ecf == '3R326', 1L, 0L),
    ecf4R401_211  = if_else(ecf %in% c('2R211', '4R401'), 1L, 0L),
    ecf4R402      = if_else(ecf == '4R402', 1L, 0L),
    ecf4R403      = if_else(ecf == '4R403', 1L, 0L),
    ecf4R404      = if_else(ecf == '4R404', 1L, 0L),
    ecf4R405      = if_else(ecf == '4R405', 1L, 0L),
    ecf4R406      = if_else(ecf == '4R406', 1L, 0L),
    ecf4R407      = if_else(ecf == '4R407', 1L, 0L),
    ecf4R408      = if_else(ecf == '4R408', 1L, 0L),
    ecf4R409      = if_else(ecf == '4R409', 1L, 0L),
    ecf4R410      = if_else(ecf == '4R410', 1L, 0L),
    ecf4R411      = if_else(ecf == '4R411', 1L, 0L)
    # (ecf5R539 was commented out in SPSS, so we omit it)
  )

# SPSS: FREQUENCIES ecf3R324 TO ecf4R411.
df |> 
  summarise(
    n_ecf3R324     = sum(ecf3R324, na.rm = TRUE),
    n_ecf3R325     = sum(ecf3R325, na.rm = TRUE),
    n_ecf3R326     = sum(ecf3R326, na.rm = TRUE),
    n_ecf4R401_211 = sum(ecf4R401_211, na.rm = TRUE),
    n_ecf4R402     = sum(ecf4R402, na.rm = TRUE),
    n_ecf4R403     = sum(ecf4R403, na.rm = TRUE),
    n_ecf4R404     = sum(ecf4R404, na.rm = TRUE),
    n_ecf4R405     = sum(ecf4R405, na.rm = TRUE),
    n_ecf4R406     = sum(ecf4R406, na.rm = TRUE),
    n_ecf4R407     = sum(ecf4R407, na.rm = TRUE),
    n_ecf4R408     = sum(ecf4R408, na.rm = TRUE),
    n_ecf4R409     = sum(ecf4R409, na.rm = TRUE),
    n_ecf4R410     = sum(ecf4R410, na.rm = TRUE),
    n_ecf4R411     = sum(ecf4R411, na.rm = TRUE)
  )
```

```{r notings2}
# 16) CONTINUED DATA EXCLUSION CHECKS (IF ANY)

# SPSS:
#   FILTER OFF.  (In R, once we filter out with dplyr::filter(), no need to re‐enable.)
```

##### CONVERT LOG PREDICTIONS TO DOLLARS

```{r log_dollars}
# SPSS:
#   COMPUTE MODMV = EXP(PRE_4).
df <- df |> 
  mutate(modmv = exp(pre_4))
```

##### Drop duplicate columns + content

```{r drop_dupe_cols}
# Suppose df is your data.frame
# 1) Transpose to get columns as rows
mat <- t(as.matrix(df))

# 2) Identify duplicated rows (i.e., duplicate columns in original)
dups <- duplicated(mat)

# 3) Keep only non-duplicates
df_clean <- df[ , !dups]

# Drop dupe column names
# df <- df[ , !duplicated(names(df))]
```

#### Some visuals

```{r vis_ggpubr}
# esp1, esp2, esp3, esp4
# ratio1, ratio2, ratio3, ratio4
ggplot(df, aes(x = s_resb_floorarea, y = saleprice)) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, color = "steelblue") +
  stat_regline_equation(
    aes(label =  paste(after_stat(eq.label), after_stat(adj.rr.label), sep = "~~~")),
    label.x.npc = "left",   # use "npc" coordinates 0–1 in panel
    label.y.npc = "top"
  )
```

##### CREATE CHARTS OF TASP AND TASPPSF BY ECF FROM THE FINAL MODEL

```{r chart_tasp, eval = FALSE}
# SPSS:
#   MEANS TASP BY ECF / CELLS=COUNT MEDIAN MEAN MIN MAX.
#   MEANS TASPPSF BY ECF / CELLS=COUNT MEDIAN MEAN MIN MAX.

df |> 
  group_by(ecf, erabuilt) |> 
  summarise(
    n               = n(),
    MEDIAN_tasp     = median(tasp, na.rm = TRUE),
    MEAN_tasp       = mean(tasp, na.rm = TRUE),
    MIN_tasp        = min(tasp, na.rm = TRUE),
    MAX_tasp        = max(tasp, na.rm = TRUE)
  )

df |> 
  group_by(ecf) |> 
  summarise(
    n              = n(),
    MEDIAN_tasppsf = median(tasppsf, na.rm = TRUE),
    MEAN_tasppsf   = mean(tasppsf, na.rm = TRUE),
    MIN_tasppsf    = min(tasppsf, na.rm = TRUE),
    MAX_tasppsf    = max(tasppsf, na.rm = TRUE)
  )
```

```{r notings3}
# 19) FINAL NOTES
# ------------------------------------------------------------------------
# At this point, all SPSS “COMPUTE” and “RECODE” statements have been translated to `mutate()` calls,
# and all SPSS “FREQUENCIES” / “MEANS ... BY ...” have been translated to `table()` or `group_by() %>%
# summarise()` calls.
#
# If you wish to save the resulting R data as an RDS, do:
#   saveRDS(df, file = 'ECF_NE-Area_transformed.rds')
#
# If you need to export to CSV:
#   write_csv(df, 'ECF_NE-Area_transformed.csv')
#
# You can also inspect any remaining variables or run further diagnostics as needed.

# END OF R TRANSLATION
```

```{r drop_nas}
df <- df |> 
  drop_na()
```

#### Regression Models

```{r reg_mod}
# 23. Regression Models (SPSS: REGRESSION /DEPENDENT=LN_tasp /METHOD=BACKWARD ...)
# The main regression model:
# Note: Ensure all variables used in the model exist in your 'df_processed' data frame.
# If some variables are not created due to previous comments in SPSS, you'll need to adjust.
# Also, R's lm() doesn't directly support backward selection as an argument.
# You'd typically use step() for stepwise selection after an initial model, or specify variables manually.

# The second regression in the file (SPSS: REGRESSION /DEPENDENT=LN_tasp /METHOD=BACKWARD ...)
# This model uses the time-adjusted sale price (LN_tasp).
# lnsbldsfINT was part of the original SPSS REGRESSION

lm_model_tasp <- lm(ln_tasp ~ ln_landratio + lnsbldsfint + arterial +
                      single_family1story + half_duplex1story + half_duplex2story +
                      smallest + small + large + largest + two_plus_baths + powder_room +
                      good + fair + poor + very_poor_unsound + qabove_avg + qbelow_avg +
                      yb_pre_1929 + yb_1946_to_1960 + yb_post_1961 + crawl + slab +
                      ln_garagespaces + fireplace + hot_water + elec_baseboard + wall +
                      central_air + months_1to9 + months_10to16 + months_17to24,
                    data = df)
summary(lm_model_tasp)

# For backward selection in R, you would typically use `step()`:
# # Define the formula for the full model
# formula_full_tasp <- as.formula(
#   "LN_tasp ~ LN_LANDRATIO + lnsbldsfINT + ARTERIAL +
#   single_family1story + half_duplex1story + half_duplex2story +
#   smallest + small + large + largest +
#   two_plus_baths + powder_room +
#   good + fair + poor + very_poor_unsound +
#   qabove_avg + qbelow_avg +
#   yb_pre_1929 + yb_1946_to_1960 + yb_post_1961 +
#   CRAWL + SLAB + LN_GARAGESPACES + FIREPLACE +
#   HOT_WATER + ELEC_BASEBOARD + WALL + CENTRAL_AIR +
#   months_1to9 + months_10to16 + months_17to24"
# )
#
# initial_lm_model_tasp <- lm(formula_full_tasp, data = df_processed)
# backward_lm_model_tasp <- step(initial_lm_model_tasp, direction = "backward", trace = FALSE) # Set trace=TRUE to see steps
# summary(backward_lm_model_tasp)

# Final descriptive statistics (SPSS: DES saleprice tasp taf tasppsf.)
# In R:
# summary(df_processed$saleprice)
# summary(df_processed$tasp)
# summary(df_processed$taf)
# summary(df_processed$tasppsf)
```

#### Clean Dataframe for machine learning input

```{r recipe}
rec <- recipe(saleprice ~ ., data = df_clean) |> 
  # 1. Remove zero‐variance predictors
  step_zv(all_predictors()) |> 
  # 2. (Optional) other preprocessing steps, e.g. centering/scaling
  step_center(all_predictors()) |> 
  step_scale(all_predictors())

```

```{r recipe2}

rec2 <- recipe(saleprice ~ ., data = df_clean) %>%
  step_zv(all_predictors()) %>%               # drop zero‐variance
  # center & scale *only* the numeric predictors
  step_center(all_numeric_predictors()) %>%
  step_scale(all_numeric_predictors())

```



```{r}
rec2 %>% tidy(number = 1)

prep_rec2 <- prep(rec2, training = df_clean)
tidy(prep_rec2, number = 1)


```



```{r}
data_split  <- initial_split(df_clean, prop = 0.8)
train_data  <- training(data_split)
test_data   <- testing(data_split)

rec_prep3    <- prep(rec2, training = train_data)
train_clean  <- bake(rec_prep3, new_data = train_data) # use to train, fit
test_clean   <- bake(rec_prep3, new_data = test_data)


```






#### Train / Test Split

```{r tt_split}
set.seed(42)

data_split <- initial_split(df4, prop = 0.8, strata = saleprice)
train_data <- training(data_split)
test_data  <- testing(data_split)
```

#### Preprocess

```{r preprocess}
house_rec <- recipe(saleprice ~ ., data = train_data) |> 
  step_impute_median(all_numeric_predictors())        |> 
  step_impute_mode(all_nominal_predictors())          |> 
  step_dummy(all_nominal_predictors())                |> 
  step_normalize(all_numeric_predictors())
```

#### Choice of Model

```{r mod_typ}
boost_spec <- boost_tree(
    trees          = tune(),
    learn_rate     = tune(),
    tree_depth     = tune(),
    loss_reduction = tune(),
    min_n          = tune()
  ) |> 
  set_engine('xgboost') |> 
  set_mode('regression')
```

##### Remove problematic columns

```{r col_del}
# df2 <- subset(df_clean, select = -s_propclass)
df4 <- subset(df_clean,
              select = -c(s_propclass, grantor, grantee, saledate, end_index,
                          is_vacant_land_sale, primary_last, elec_baseboard, rate2, rate3,
                          yb_pre_1910))

# You can remove all columns between d and b with:
# Data <- subset( Data, select = -c( d : b )

# step_zv()
```



#### Training?

```{r train_mod}
library(workflows)
wf <- workflow() %>%
  add_recipe(rec) %>%
  add_model(your_spec)

# Fit on the training slice
fit_wf <- fit(wf, data = train_data)

# Predict on test slice
predict(fit_wf, test_data)

```



#### Tune hyper‑parameters

```{r tune_hp}
set.seed(42)

grid <- grid_latin_hypercube(
  trees(), learn_rate(), tree_depth(),
  loss_reduction(), min_n(),
  size = 20
)

folds <- vfold_cv(train_data, v = 5)

wf <- workflow() |> 
  add_recipe(house_rec) |> 
  add_model(boost_spec)

tuned <- tune_grid(
  wf, resamples = folds, grid = grid,
  metrics = metric_set(rmse, rsq, mae),
  control = control_grid(save_pred = TRUE)
)

best_params <- select_best(tuned, metric = 'rmse') # show_best(tuned, metric = "rmse")

final_wf <- finalize_workflow(wf, best_params) |> 
  fit(train_data)

# show_notes(.Last.tune.result)
```


#### Evaluate Model

```{r eval_mod}
preds <- predict(final_wf, test_data) |> 
  bind_cols(test_data)

metrics <- preds |> 
  metrics(truth = sale_price, estimate = .pred)

metrics
#>       .metric .estimator      .estimate
#> 1        rmse      standard   18234.19
#> 2         rsq      standard       0.82
#> 3         mae      standard   13571.47
```

#### Save and Reuse

```{r sav_mod}
bundled <- bundle(final_wf) |>
  axe_call()
saveRDS(bundled, "detroit_house_price_model.rds")
```

#### Predict a new listing

```{r mak_predict}
new_house <- tibble(
  bedrooms        = 3,
  baths_tot       = 1.5,
  finished_sq_ft  = 1200,
  age             = 75,
  stories         = 2,
  bldg_cond       = "Average",
  bldg_mat        = "Brick",
  fireplace       = 1,
  garage          = 1
)

model     <- readRDS("detroit_house_price_model.rds") |>
  unbundle()

predicted <- predict(model, new_house)

scales::dollar(predicted$.pred)
#> $148,000

```





























